---
title: 'Clean Discharge & Sampling Data'
subtitle: 'Tasks: Flowmeter Error Removal'
author: "PAZ"
date: "25 octobre 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

## Purpose

This document removes aberrant discahrge values in the flow meter data. 

Used files:

1. **Alteck2016Debit.csv**

Produced file:

1. **hydroAlteck2016_NAs_R.csv** (Used for testing methods to replace missing values and smoothing the time series).

## Required R-packages:

```{r, message=FALSE}

# Plotting functions
library("ggplot2")
library("scales")
library("tidyr")

```

## Working directory

```{r, message=FALSE}

# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")
# setwd("/Users/DayTightChunks/Documents/PhD/Routput/Alteck/R")
# setwd("D:/Documents/these_pablo/Alteckendorf2016/00_TransparencyFolder/Discharge")
getwd()

```

## Import raw discharge data

```{r, message=FALSE}

dischargeAlteck = read.csv2("Data/Alteck2016Debit.csv", header = FALSE)

dischargeAlteck$Date = as.POSIXct(strptime(dischargeAlteck$V1, 
                                           "%d/%m/%Y %H:%M", 
                                           tz="EST"))
sum(is.na(dischargeAlteck$Date))

dischargeAlteck <- dischargeAlteck[!duplicated(dischargeAlteck$V1),]

dischargeAlteck = dischargeAlteck[,c(3,1:2)]
colnames(dischargeAlteck) <- c("Date", "DateCheck", "Q.m3Hrs")
head(dischargeAlteck)

```


## Replace aberrant values with NA's

To evaluate which data needed correction and overview of the data was necessary:

```{r, message=FALSE}

altp <- ggplot(dischargeAlteck, aes(x=Date, y=Q.m3Hrs))
altp + geom_line(colour = "blue") +
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h))))
```

First, negative values were replaced with NA:

```{r, message=FALSE}

dischargeAlteck$Qna = dischargeAlteck$Q.m3Hrs
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs <= 0.0] = NA

```

A number of subsections were then manually selected and unrealistic values converted to NA entries. Note that past May 30, only negative values have been removed (replaced by NA).

```{r, message=FALSE}


dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 1.0 & 
                      dischargeAlteck$Date > as.POSIXct("2016-03-29 23:00:00 EST") & 
                      dischargeAlteck$Date < as.POSIXct("2016-03-31 00:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 6 & 
                      dischargeAlteck$Date > as.POSIXct("2016-04-01 23:00:00 EST") & 
                      dischargeAlteck$Date < as.POSIXct("2016-04-04 00:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 7.5 & 
                      dischargeAlteck$Date > as.POSIXct("2016-04-06 23:00:00 EST") & 
                      dischargeAlteck$Date < as.POSIXct("2016-04-07 15:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 2.5 & 
                      dischargeAlteck$Date > as.POSIXct("2016-04-9 00:00:00 EST") & 
                      dischargeAlteck$Date < as.POSIXct("2016-04-17 00:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 2.5 & 
                      dischargeAlteck$Date > as.POSIXct("2016-05-07 23:00:00 EST") & 
                      dischargeAlteck$Date < as.POSIXct("2016-05-09 23:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 3.0 & 
                      dischargeAlteck$Date > as.POSIXct("2016-05-22 23:00:00 EST") &
                      dischargeAlteck$Date < as.POSIXct("2016-05-30 23:00:00 EST") ] = NA
dischargeAlteck$Qna[dischargeAlteck$Q.m3Hrs < 100.0 & 
                      dischargeAlteck$Date > as.POSIXct("2016-05-13 06:00:00 EST") &
                      dischargeAlteck$Date < as.POSIXct("2016-05-13 14:00:00 EST") ] = NA



```


The following is an example of the removed abberant values:

1. Example 1

Red geom_line has been omitted in final data. 

```{r, message=FALSE}

altpTest = ggplot(dischargeAlteck, aes(x=Date)) + 
  geom_line(aes(y= Q.m3Hrs), colour = "red") +
  geom_line(aes(y=Qna), colour = "blue") +
  coord_cartesian(xlim = c(as.POSIXct("2016-06-11 23:00:00 EST"), 
                           as.POSIXct("2016-07-12 23:00:00 EST")), 
                  ylim = c(0, 100)) + # no.1
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  #scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h))))

altpTest


```

2. Example 2

Red geom_line has been ommitted infinal data. 

```{r, message=FALSE}

altpTest = ggplot(dischargeAlteck, aes(x=Date)) +
  geom_line(aes(y= Q.m3Hrs), colour = "red") +
  geom_line(aes(y= Qna), colour = "blue") +
  coord_cartesian(xlim = c(as.POSIXct("2016-05-13 06:00:00 EST"), 
                           as.POSIXct("2016-05-13 14:00:00 EST"))
                  , 
                  ylim = c(0, 100)
                  ) +
  theme_bw() +
  #scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  #scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h))))

altpTest

head(dischargeAlteck)

```



## Save files

```{r}

head(dischargeAlteck)
write.csv2(dischargeAlteck, "Data/hydroAlteck2016_NAs_R.csv", row.names = FALSE)

```

