---
title: "Statistical Modelling of Missing Discharge Data"
author: "PAZ"
date: "25 octobre 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)

Sys.setlocale("LC_ALL", "English")
```


## Purpose

This document compares three methods to:

Input missing data in cleaned discharge data:

- Simple interpolation (inputting data)

Smooth-out the time series.
- Exponential Weighted Moving Averages (EWMA) - i.e. parameter alpha is adjusted manually
- Double Exponential Smoothing - i.e. R finds optimal parameters automatically

The input file is: 

- **hydroAlteck2016_NAs_R.csv**

The file stems from *CleanDischargeDat_hydroAlteck2016_NAs.Rmd*, which removed aberrant values from the flow meter data.

The generated output file is:  

- **hydroAlteck2016_smooth_R.csv**. 


## Required packages

```{r, message=FALSE}

# Plotting functions
library("ggplot2")
library("scales")
library("tidyr")

# Interpolation packages
library("zoo")
library("forecast")

```

## Import "clean" discharge data

```{r}
hydroAlteck2016_na = read.csv2("Data/hydroAlteck2016_NAs_R.csv")
hydroAlteck2016_na$Date = as.POSIXct(strptime(hydroAlteck2016_na$Date, "%Y-%m-%d %H:%M", tz="CET"))

head(hydroAlteck2016_na)

altp <- ggplot(hydroAlteck2016_na, aes(x=Date, y=Qna))
altp <- altp + geom_line(colour = "blue") +
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h)))) +
  scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000))

altp

# + coord_cartesian(xlim = c(as.POSIXct("2016-05-08 23:00:00 CET"), 
#                           as.POSIXct("2016-07-12 23:00:00 CET"))
#                  , ylim = c(0, 100)
#                  ) # no.1
  #scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m"))

```


## 1st Discharge Set - Approximating Missing Data via the **Zoo** package

The **Zoo** package is one of the few packages (i.e. also **forecast**) where inputing data to univariate time series is possible [@Moritz2015]. Functions include:

- na.aggregate()
- na.StructTS()
- na.locf()
- na.approx()
- na.spline()

## na.approx() function

Missing values (NAs) are replaced by linear interpolation using the na.approx function.

```{r}
hydroAlteck2016_na$Qna[nrow(hydroAlteck2016_na)] = 0
  hydroAlteck2016_na$Qapprox = na.approx(hydroAlteck2016_na$Qna)
  head(hydroAlteck2016_na)

interpol <- ggplot(hydroAlteck2016_na, aes(Date)) + 
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h)))) +
  scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000)) +
  geom_line(aes(y = hydroAlteck2016_na$Qapprox), color="red") +
  geom_line(aes(y = hydroAlteck2016_na$Qna), color="blue") # + 
  # coord_cartesian(xlim = c(as.POSIXct("2016-03-29 23:00:00 CET"), as.POSIXct("2016-04-05 00:00:00 CET"))
  #                  , ylim = c(0, 100)
  #                  )  
interpol


                
```

## na.StructTS() function (not working, can't convert to ts object with freq.)

```{r}

# Code for na.StructTS
                
```

## na.interp() function 

This function shows no improvement over the na.approx() method.

```{r}


hydroAlteck2016_na$Qinterp = na.interp(hydroAlteck2016_na$Qna)
head(hydroAlteck2016_na)

interpol <- ggplot(hydroAlteck2016_na, aes(Date)) + 
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h)))) +
  scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000)) +
  geom_line(aes(y = hydroAlteck2016_na$Qapprox), color="red") +
  geom_line(aes(y = hydroAlteck2016_na$Qinterp), color="forestgreen") +
  geom_line(aes(y = hydroAlteck2016_na$Qna), color="blue") #  + 
  # coord_cartesian(xlim = c(as.POSIXct("2016-03-29 23:00:00 CET"), as.POSIXct("2016-04-05 00:00:00 CET"))
  #                  , ylim = c(0, 100)
  #                  )  
interpol

                
```


## Smoothing Data 

### Holt Winters 1 - Exponential Weighted Moving Averages (EWMA)

This approach manually adjusts the value of alpha.

```{r}

# plot.ts(hydroAlteck2016_na$Qinter)
Q.HW1mean <- HoltWinters(hydroAlteck2016_na$Qinter, 
                       alpha = 0.2, # If larger, less damping (i.e. more reactive).
                       beta = FALSE, # Controls how the trend adapts
                       gamma = FALSE # Controls adaptation of seasonal values
                       ) 

# Note:
# beta=False and gamma=FALSE gives Exponential Weighted Moving Averages (EWMA)

Q.HW1mean

# Removing the first entry of the original data to merge model
hydroAlteck2016 = hydroAlteck2016_na[2:nrow(hydroAlteck2016_na),]
hydroAlteck2016$Q.HW1 = Q.HW1mean$fitted[,1]

head(hydroAlteck2016)

```


### Holt Winters 2 - Double Exponential Smoothing

This approach manually adjusts the value of alpha.

```{r}

Q.HW2mean <- HoltWinters(hydroAlteck2016_na$Qinter, 
                        gamma = FALSE) 

Q.HW2mean
# Note:
# gamma=FALSE gives Double Exponential Smoothing

# Shorten the data set by one more observation
hydroAlteck2016 = hydroAlteck2016[2:nrow(hydroAlteck2016),]
hydroAlteck2016$Q.HW2 = Q.HW2mean$fitted[,1]

head(hydroAlteck2016)

```


### Plotting the two smoothing methods

```{r}
Qsmooth <- ggplot(hydroAlteck2016, aes(Date)) + 
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h)))) +
  scale_y_continuous(trans=log_trans(), breaks=c(1,10, 50, 100, 500,1000)) +
  geom_line(aes(y = hydroAlteck2016$Qinter), color="black") +
  geom_line(aes(y = hydroAlteck2016$Q.HW2), color="forestgreen") +
  geom_line(aes(y = hydroAlteck2016$Q.HW1), color="blue")  + 
   coord_cartesian(xlim = c(as.POSIXct("2016-04-10 23:00:00 CET"), as.POSIXct("2016-04-15 00:00:00 CET"))
  #                  , ylim = c(0, 100)
                    )  
Qsmooth
```


## Approximating missing values via subset prediction (trends)

This section needs to subset the missing data and treated separetely.

```{r}

Q1.predict <- predict(Q.HW1mean, 
                      n.ahead = 10,
                      prediction.interval = TRUE)

Q1.predict
# Q1.mean$fitted
plot.ts(log(hydroAlteck2016_na$Qinter))
lines(log(Q.HW2mean$fitted[,1]), col="green")
lines(log(Q.HW1mean$fitted[,1]), col="blue")
```



## Approximating Missing Data - Local Level Model

The local level model assumes that we observe a time series, $y_t$, and that time series is the sum of another time series, $\mu_t$, and random,
corrupting noise, $e_t$. We would prefer to directly observe $\mu_t$, a latent variable, but cannot due to the noise.

### Establish the model

```{r}
struct1 <- StructTS(hydroAlteck2016_na$Qinter, type="level")
if (struct1$code != 0) stop("optimizer did not converge")
print(struct1$coef)

cat("Transitional variance:", struct1$coef["level"],
    "\n", "Observational variance:", struct1$coef["epsilon"],
    "\n", "Initial level:", struct1$model0$a, "\n")
```

### Filter the with the StrucTS Model created

```{r}
filt <- KalmanRun(hydroAlteck2016_na$Qinter, struct1$model)
#plot(unlist(filt))
```

Stuck trying to filter the data based on the model... skipping now.

# Variable to Use
```{r}
altp <- ggplot(hydroAlteck2016, aes(x=Date, y=Q.HW1))
altp <- altp + geom_line(colour = "blue") +
  theme_bw() +
  scale_x_datetime(breaks = date_breaks("weeks"), labels = date_format("%d/%m")) +
  theme(axis.text.x=element_text(angle = 45, hjust = 0.75)) +
  xlab("") +
  ylab(expression(paste("Q ",({m}^"3"/h)))) +
  scale_y_continuous(trans=log_trans(), breaks=c(1,10,100,1000))

altp
```



## Saving

```{r}

write.csv2(hydroAlteck2016, "Data/hydroAlteck2016_smooth_R.csv", row.names = FALSE)

```





