---
title: "Sub-weekly Discharge - Alteck 2016, Part I"
author: "PAZ"
date: "25 octobre 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

## Purpose

This file computes average discharge and total volume discharged per sub-week (sampled and non-sampled).

Imports:

- **Alteck2016Debit_smooth_R.csv** 

Generates: 

- **WeeklyHydro_R.csv** 

which is a file with discharge data summarized by sub-week. 

Columns in **WeeklyHydro_R.csv** include:

1. Week-SubWeek ID
2. Average Discharge
3. Volumne discharged
4. Elapsed hours per subweek
5. Marker indicating whether the subweek was sampled or not.

To see the variables in this file see the end of this document. 

## Required R-packages:

```{r, message=FALSE}

# Date-time functions
library(chron)
```

## Working directory

```{r, message=FALSE}

# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")
# setwd("/Users/DayTightChunks/Documents/PhD/Routput/Alteck/R")
getwd()

```

## Import smooth (and sampler merged) discharge data

```{r, message=FALSE}

debitAlt = read.csv2("Data/hydroAlteck2016_R.csv", header = TRUE)
head(debitAlt)
tail(debitAlt)
sum(is.na(debitAlt$Q.HW1))
```

## Convert Date column to a readable Date-time object

```{r, message=FALSE}

class(debitAlt$Date)
debitAlt$Date = as.POSIXct(strptime(debitAlt$Date, "%Y-%m-%d %H:%M", tz="EST"))
sum(is.na(debitAlt$Date))
sum(is.na(debitAlt$Q.HW1))

```

## Choose the variables to report as sub-weekly summaries

1. Discharge (m3/h) variable to calculate sub-weekly averages
2. Rainfall (mm) taken every 2 minutes, will become cummulative for the sub-weekly sample

```{r, message=FALSE}

# Remove all unecessary columns
debitAlt = debitAlt[,c("Date", "Q.HW1", "Rain.mm")]
head(debitAlt)
```


## Define the sub-week sample dates

During the 2016 campaign some periods where not sampled because the automatic sampler either experienced malfunction or the capacity was reached before intervention. Interventioned took place on a weekly basis. 

1. Extract subsets where sampling was conducted.
2. Extract subsets where sampling was not conducted.

```{r, message=FALSE}

# Define the Weekly discharge subsets
W00 <- subset(debitAlt, 
              Date < as.POSIXct('2016-03-25 12:04:00' , tz="EST") ) # Not sampled

W01 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-03-25 12:04:00' , tz="EST") & 
                Date < as.POSIXct('2016-03-28 22:37:00' , tz="EST"))

W02 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-03-28 22:37:00' , tz="EST") & 
                Date <= as.POSIXct('2016-03-30 12:17:00' , tz="EST")) # Not sampled

W11 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-03-30 12:17:00' , tz="EST") & 
                Date <= as.POSIXct('2016-03-31 15:35:00' , tz="EST"))
W12 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-03-31 15:35:00' , tz="EST") & 
                Date <= as.POSIXct('2016-04-01 14:45:00' , tz="EST"))
# W13
W13 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-01 14:45:00' , tz="EST") & 
                Date <= as.POSIXct('2016-04-05 15:07:00' , tz="EST")) # Not sampled

W21 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-05 15:07:00' , tz="EST") & 
                Date <= as.POSIXct('2016-04-06 14:51:00' , tz="EST"))
W22 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-06 14:51:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-08 00:38:00' , tz="EST"))

# W23 not sampled
W23 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-08 00:38:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-14 13:51:00' , tz="EST"))

W31 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-14 13:51:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-16 18:32:00' , tz="EST"))

W32 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-16 18:32:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-17 09:02:00' , tz="EST"))
###
W32.1 <- subset(debitAlt, 
                Date >= as.POSIXct('2016-04-17 09:02:00' , tz="EST") & 
                  Date < as.POSIXct('2016-04-18 20:30:00' , tz="EST")) #  Not sampled

W33 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-18 20:30:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-21 09:11:00' , tz="EST"))

W41 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-21 09:11:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-23 06:37:00' , tz="EST"))

W42 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-23 06:37:00' , tz="EST") & 
                Date < as.POSIXct('2016-04-26 11:50:00' , tz="EST")) # Not sampled

W51 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-04-26 11:50:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-01 10:46:00' , tz="EST"))

W52 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-01 10:46:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-03 12:02:00' , tz="EST"))

##
W53 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-03 12:02:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-03 13:09:00' , tz="EST")) # Not sampled

W61 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-03 13:09:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-10 00:05:00' , tz="EST"))

W62 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-10 00:05:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-12 06:33:00' , tz="EST"))
W63 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-12 06:33:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-12 09:12:00' , tz="EST"))
W64 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-12 09:12:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-12 12:52:00' , tz="EST"))

##
W65 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-12 12:52:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-13 12:05:00' , tz="EST")) # Not sampled

W71 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-13 12:05:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-16 15:11:00' , tz="EST"))
##
W72 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-16 15:11:00' , tz="EST") & 
                Date <= as.POSIXct('2016-05-17 09:16:00' , tz="EST")) # Not sampled

W81 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-17 09:16:00' , tz="EST") & 
                Date <= as.POSIXct('2016-05-23 18:02:00' , tz="EST"))
##
W82 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-23 18:02:00' , tz="EST") & 
                Date <= as.POSIXct('2016-05-24 12:00:00' , tz="EST")) # Not sampled

W91 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-24 12:00:00' , tz="EST") & 
                Date <= as.POSIXct('2016-05-29 12:09:00' , tz="EST"))
W92 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-29 12:09:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-30 05:48:00' , tz="EST"))
W93 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-30 05:48:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-30 12:11:00' , tz="EST"))
W94 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-30 12:11:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-30 17:28:00' , tz="EST"))
##
W95 <- subset(debitAlt, 
              Date >= as.POSIXct('2016-05-30 17:28:00' , tz="EST") & 
                Date < as.POSIXct('2016-05-31 12:00:00' , tz="EST")) # Not sampled

W101 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-05-31 12:00:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-02 12:57:00' , tz="EST"))
W102 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-02 12:57:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-03 12:05:00' , tz="EST"))
W103 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-03 12:05:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-04 08:35:00' , tz="EST"))
W104 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-04 08:35:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-04 11:00:00' , tz="EST"))
W105 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-04 11:00:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-04 15:31:00' , tz="EST"))
#
W106 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-04 15:31:00' , tz="EST") & 
                 Date <= as.POSIXct('2016-06-07 12:00:00' , tz="EST")) # Not sampled

W111 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-07 12:00:00' , tz="EST") & 
                 Date <= as.POSIXct('2016-06-10 05:25:00' , tz="EST"))
W112 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-10 05:25:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-14 12:34:00' , tz="EST"))
W113 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-14 12:34:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-14 13:06:00' , tz="EST")) 
W121 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-14 13:06:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-15 08:14:00' , tz="EST")) 
W122 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-15 08:14:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-16 08:21:00' , tz="EST"))
W123 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-16 08:21:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-17 00:49:00' , tz="EST"))

W124 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-17 00:49:00' , tz="EST") & 
                 Date <= as.POSIXct('2016-06-17 11:05:00' , tz="EST"))
#
W125 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-17 00:49:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-21 12:00:00' , tz="EST")) # Not sampled

W131 <- subset(debitAlt, 
               Date >= as.POSIXct('2016-06-21 12:00:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-24 14:51:00' , tz="EST"))
W132 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-24 14:51:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-25 07:49:00' , tz="EST"))
W133 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-25 07:49:00' , tz="EST") & 
                 Date < as.POSIXct('2016-06-28 08:55:00' , tz="EST"))
W141 <- subset(debitAlt, 
               Date > as.POSIXct('2016-06-28 08:55:00' , tz="EST") & 
                 Date <= as.POSIXct('2016-07-04 14:41:00' , tz="EST"))

W151 <- subset(debitAlt, 
               Date > as.POSIXct('2016-07-04 14:41:00' , tz="EST") & 
                 Date <= as.POSIXct('2016-07-12 10:20:00' , tz="EST"))

```

## Calculate sampled & non-sampled hrs.

1. Create a list made up of each subset data.frame

```{r, message=FALSE}

# Create a list of matrices

weeks = list(W01, 
             W11, 
             W12, 
             W21, W22, 
             W31, W32, W33, 
             W41, 
             W51, W52, 
             W61, W62, W63, W64, 
             W71, 
             W81, 
             W91, W92, W93, W94,
             W101, W102, W103, W104, W105,
             W111, W112, W113,
             W121, W122, W123, W124,
             W131, W132, W133,
             W141,
             W151)

weeksNS = list(W00, W02, W13, W21, W32.1, W42, W53, W65, W72, W82, W95, W106, W125)

```


2. Create a function to calculate the time difference between the last and first date-entry of each subset.

```{r, message=FALSE}

hoursInWeek = function(x){
  as.numeric(difftime(x[nrow(x),1], x[1,1], units = "hours"), units = "hours")
} 

```

Here "x" is a matrix (i.e. data.frame object) with number of rows = nrow. The *difftime* function uses the last index-value (t2) - first index-value (t1)

The *difftime* syntax: 
matrix[last_row, first_column], matrix[first_row, first_column] -> t2 - t1 

3. Get total hours for each subset (data.frame) with the *sapply* function 

```{r, message=FALSE}
sampledHrsList = sapply(weeks, hoursInWeek)
nonSampledHrsList = sapply(weeksNS, hoursInWeek)

nonSampledHrsList

```

Here, the function *sapply* returns a list of outputs based on the created function we pass to it. The syntax: sapply(object, function to apply on each object).

4. Get volume discharged. 

Same as above, we create first a function and then use *sapply*. Note that the discharge data in [$m^3 h^{-1}$] has an interval of 2 minutes. So we multiply each entry of the discharge column by 2 and convert to hrs. 
 
```{r, message=FALSE}

# m^3/h * 2 min * 1h/60min = m3
volMinute = function(x) {
  (x[,2] * 2/60)
}

volMinuteList = sapply(weeks, volMinute)
volMinuteListNS = sapply(weeksNS, volMinute)

```

5. Get total rainfall per sub-week

```{r}
rainTot = function(x) {
  sum(x[,3])
}

rainList = sapply(weeks, rainTot)
rainListNS = sapply(weeksNS, rainTot)
```



**Check results**

```{r, message=FALSE}

# first matrix in the weeks list (note: matix has two rows)
# weeks[[1]] 

# access the first matrix in the list, the second element in THAT list.
volMinuteList[[1]][[2]] 

# Cumm rain per sub-sample
rainList
```


6. Get total volume for each subset (i.e. each subweek)

```{r, message=FALSE}

total = function(x){
  sum(x)
}

volumesList = sapply(volMinuteList, total)
volumesListNS = sapply(volMinuteListNS, total)

# Check volumnes for the non-sample list:
volumesListNS

```

7. Create an average discharge and an average rainfall intensity list for each entry result above

```{r, message=FALSE}

AveDischarge = volumesList/sampledHrsList
AveDischargeNS = volumesListNS/nonSampledHrsList

RainIntensity = rainList/sampledHrsList
RainIntensityNS = rainListNS/nonSampledHrsList

```

8. Put all data together into a new data.frame structure.

```{r, message=FALSE}

# Sampled data frame

ID = c("W0-1", "W1-1", "W1-2", 
       "W2-1", "W2-2", 
       "W3-1", "W3-2", "W3-3", "W4-1", 
       "W5-1", "W5-2", 
       "W6-1", "W6-2", "W6-3", "W6-4", 
       "W7-1", 
       "W8-1", 
       "W9-1", "W9-2", "W9-3", "W9-4",
       "W10-1", "W10-2", "W10-3", "W10-4", "W10-5",
       "W11-1", "W11-2", "W11-3",
       "W12-1", "W12-2", "W12-3", "W12-4",
       "W13-1", "W13-2", "W13-3",
       "W14-1",
       "W15-1")

Sampled = rep("Sampled", length(ID))


WeeklyDischarge = as.data.frame(ID)
WeeklyDischarge$AveDischarge.m3.h = AveDischarge
WeeklyDischarge$Volume.m3 = volumesList
WeeklyDischarge$Sampled.Hrs = sampledHrsList
WeeklyDischarge$Sampled = Sampled
WeeklyDischarge$CumRain.mm = rainList
WeeklyDischarge$RainInt.mmhr = RainIntensity

# Not sampled data frame

IDns = c("W0-0x", "W0-2x", "W1-3x", "W2-3x", "W3-2.1x", "W4-2x", 
         "W5-3x", "W6-5x", "W7-2x", "W8-2x", 
         "W9-5x", "W10-6x", "W12-5x")

SampledNS = rep("Not Sampled", length(IDns))

WeeklyDischargeNS = as.data.frame(IDns)
WeeklyDischargeNS$AveDischarge.m3.h = AveDischargeNS
WeeklyDischargeNS$Volume.m3 = volumesListNS
WeeklyDischargeNS$Sampled.Hrs = nonSampledHrsList
WeeklyDischargeNS$Sampled = SampledNS
WeeklyDischargeNS$CumRain.mm = rainListNS
WeeklyDischargeNS$RainInt.mmhr = RainIntensityNS

colnames(WeeklyDischargeNS)[1] <- "ID"

# Bind both data frames

WeeklyHydro = rbind(WeeklyDischarge, WeeklyDischargeNS)
WeeklyHydro$ID = as.character(WeeklyHydro$ID)
WeeklyHydro = with(WeeklyHydro, WeeklyHydro[order(ID),])

head(WeeklyHydro)

```


## Save files

```{r, message=FALSE}

VolNS = sum(WeeklyDischargeNS$Volume.m3)
VolS = sum(WeeklyDischarge$Volume.m3)
print("% Not sampled: ")
(VolNS/(VolNS+VolS))*100  

write.csv2(WeeklyHydro, 
           'Data/WeeklyHydro_R.csv', row.names = F)
```

