---
title: "Mass Discharge - Outlet Alteck. 2016"
author: "PAZ"
date: "27 octobre 2016"
output: pdf_document
---


```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

## Purpose

This file computes the discharged mass observed at the outlet. To do that it imports the weekly discharge summary and lab results for isotopes ($^{13}C$) and s-metolachlor concentrations.

Imports: 

- **WeeklyHydro_R.csv** 
- **fluxAlteck2016_R.csv**
- **OutletConc_W0toW17.csv**
- **MESAlteckWater.csv**
- **Outlet_Isotopes_W0toW17.csv**
- **MESAlteck_FilterIsotopes.csv**

- **AO-Hydrochem.csv**

Generates:

- **WeeklyHydroContam_R.csv**

## Required R-packages:

```{r, message=FALSE}

library("plyr")
library("dplyr")

```

## Working directory

```{r, message=FALSE}

# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")
# setwd("/Users/DayTightChunks/Documents/PhD/Routput/Alteck/R")
# setwd("D:/Documents/these_pablo/Alteckendorf2016/00_TransparencyFolder")
getwd()

```


## Outlet Data - Alteckendorf 2016

1. Hydrological data on a subweekly basis

```{r, message=FALSE}

weeklyhydro = read.csv2("Data/WeeklyHydro_R.csv", header = TRUE)
colnames(weeklyhydro)[colnames(weeklyhydro) == "ID"] <- "WeekSubWeek"
head(weeklyhydro)

weeklyflux = read.csv2("Data/fluxAlteck2016_R.csv", header = TRUE)
head(weeklyflux)
```

2. Concentration data (dissolved and suspended solids)

```{r, message=FALSE}

outletConc = read.csv2("Data/OutletConc_W0toW17.csv", header = T)
outletConc <- outletConc[outletConc$ID4 != "J+7", ]
outletConc <- outletConc[,c("WeekSubWeek", "Conc.mug.L", "Conc.SD")]
head(outletConc)

filters = read.csv2("Data/MESAlteckWater.csv")
filters$MO.mg.L = ifelse(filters$MO.mg.L < 0, 0.0001, filters$MO.mg.L)
head(filters)
```

3. Isotope data

```{r, message=FALSE}

# Outlet isotope data:
outletIso = read.csv2("Data/Outlet_Isotopes_W0toW17.csv", header = T)
head(outletIso)

# Filter isotope data:
filtersIso = read.csv2("Data/MESAlteck_FilterIsotopes.csv", header = T)
filtersIso$WeekSubWeek = paste(filtersIso$Week, filtersIso$Num, sep = "-")
filtersIso <- filtersIso[filtersIso$Levl != "J+7", ]
head(filtersIso)



```

4. Hydrochemistry Data

```{r, message=FALSE}

hydroChem = read.csv2("Data/AO-Hydrochem.csv", header = T)
hydroChem = hydroChem[, c("WeekSubWeek", 
                          "NH4.mM", 
                          "TIC.ppm.filt", 
                          "Cl.mM", 
                          "NO3...mM", 
                          "PO4..mM", 
                          "NPOC.ppm" ,
                          "TIC.ppm.unfilt", 
                          "TOC.ppm.unfilt" )]
head(hydroChem)


```


## Summarizing IRMS data

```{r, message=FALSE}

isoOutSummary = ddply(outletIso, c("WeekSubWeek"), summarise,
                         N    = length(d.13C.12C),
                         diss.d13C = mean(d.13C.12C),
                         SD.d13C = sd(d.13C.12C),
                         se.d13C = SD.d13C / sqrt(N))

head(isoOutSummary)


isoFiltSummary = ddply(filtersIso, c("WeekSubWeek"), summarise,
                         N    = length(d.13C.12C),
                         filt.d13C = mean(d.13C.12C),
                         filt.SD.d13C = sd(d.13C.12C),
                         filt.se.d13C = filt.SD.d13C / sqrt(N))
head(isoFiltSummary)
```


## Merging and data wrangling stepts

1. Merge all data sets by the *WeekSubWeek* column ID, icluding:

```{r}

# Dissolved
out.CoIs = merge(outletConc, isoOutSummary, by = "WeekSubWeek", all = T)

# Filters (MES, Conc.MES)
out.CoIs = merge(out.CoIs, filters, by = "WeekSubWeek", all = T)
out.CoIs = merge(out.CoIs, isoFiltSummary, by= "WeekSubWeek", all = T)

# Pure and cuve isotope average
d13Co = -31.21

# Lab enrichment:
# epsilon = -1.61

# Lab enrichment:
# Alteck
epsilon_max = -1.5 # +/- 0.3 (@ 20C, 20% vwc)
epsilon_min = -2.0 # +/- 0.2 (@ 20C, 40% vwc)
epsilon_mean = -1.75

# Remaining fraction
out.CoIs$DD13C.diss <- (out.CoIs$diss.d13C - (d13Co))
out.CoIs$DD13C.filt <- (out.CoIs$filt.d13C - (d13Co))

out.CoIs$f.diss <- (((10**(-3)*out.CoIs$diss.d13C + 1)/(10**(-3)*d13Co + 1))**(1000/(epsilon_mean)))
out.CoIs$f.filt <- (((10**(-3)*out.CoIs$filt.d13C + 1)/(10**(-3)*d13Co + 1))**(1000/(epsilon_mean)))

out.CoIs$B.diss <- (1 - out.CoIs$f.diss)*100 
out.CoIs$B.filt <- (1 - out.CoIs$f.filt)*100 
#out.CoIs$invf <- 1/out.CoIs$f 

# Discharge times
out.CoIs = merge(weeklyhydro, out.CoIs, by = "WeekSubWeek", all = T)

# Discharge summary
out.CoIs = merge(weeklyflux, out.CoIs, by = "WeekSubWeek", all = T)

# Hydrochemistrty
out.CoIs = merge(out.CoIs, hydroChem, by= "WeekSubWeek", all = T)


out.CoIs$tf <- as.POSIXct(out.CoIs$tf, "%Y-%m-%d %H:%M", tz = "EST")
out.CoIs$ti <- as.POSIXct(out.CoIs$ti, "%Y-%m-%d %H:%M", tz = "EST")
class(out.CoIs$tf)
sum(is.na(out.CoIs$tf))

# Temprarily remove Weeks 16 & 17 (need to get discharge data)
out.CoIs <- out.CoIs[!is.na(out.CoIs$tf), ]

```

2. Weekly Exported Solids (Kg)

```{r}
# V[m3] * MES [mg/L] * 1000 [L/m3] * [1 Kg/10^6 mg]
out.CoIs$ExpMES.Kg = out.CoIs$Volume.m3*out.CoIs$MES.mg.L/1000
```

3. Weekly exported S-metolachlor mass (mg)

```{r}

# Dissolved - [mg] S-metolachlor exported per sub-week
# Conc. [mu.g s-meto/L H20] * Vol[m3] * [10^3 L/m^3] * [1 mg/10^3 mu.g]
out.CoIs$DissSmeto.mg = out.CoIs$Conc.mug.L*out.CoIs$Volume.m3  

# Solids - [mg] S-metolachlor in solids exported per sub-week 
# Conc. [mu.g s-meto / g MES] * Kg MES * [10^3 g/Kg] * [1 mg/10^3 mu.g]
out.CoIs$FiltSmeto.mg = out.CoIs$Conc.Solids.mug.gMES*out.CoIs$ExpMES.Kg 

# Total
out.CoIs$TotMassOut.mg = out.CoIs$DissSmeto.mg + out.CoIs$FiltSmeto.mg


# Proportion in dissolved and suspended solids
out.CoIs$FracDiss = out.CoIs$DissSmeto.mg/out.CoIs$TotMassOut.mg
out.CoIs$FracFilt = out.CoIs$FiltSmeto.mg/out.CoIs$TotMassOut.mg

```

4. Add the application dates and merge the total mass to the nearest discharge event

The five application dates were:

- 2016-03-20 
- 2016-04-05 
- 2016-04-13 and 2016-04-14 
- 2016-05-26 

So the total applied mass mass is merged at the nearest sampling time marker available : 

```{r}

ti = c(as.POSIXct('2016-03-25 00:04:00' , tz="EST"),
       as.POSIXct('2016-04-05 15:08:00' , tz="EST"),
       as.POSIXct('2016-04-14 13:52:00' , tz="EST"),
       as.POSIXct('2016-05-24 12:00:00' , tz="EST"))

Appl.Mass.g = c(6369.396, 3128.475, 4744.571, 4982.038)

applics = as.data.frame(ti)
applics$Appl.Mass.g = Appl.Mass.g

out.CoIs = merge(out.CoIs, applics, by = "ti", all = T)
out.CoIs$Appl.Mass.g <- ifelse(is.na(out.CoIs$Appl.Mass.g), 0.0, out.CoIs$Appl.Mass.g)

```

5. This section converts the observed S-metolachlor concentrations to [g] in dissolved water and suspended solids, assuming 0 for the values where no sample was taken. An approximative model will be tested at a later stage. 

```{r}

# Cumulative IN
out.CoIs$CumAppMass.g = cumsum(out.CoIs$Appl.Mass.g)

# First simulate a mass out to deal with missing values
# Option 1, just assume 0.0 
out.CoIs$SimOutDiss.g = out.CoIs$DissSmeto.mg/10^3
out.CoIs$SimOutFilt.g = out.CoIs$FiltSmeto.mg/10^3
out.CoIs$SimOutDiss.g = ifelse(is.na(out.CoIs$SimOutDiss.g), 0.0, out.CoIs$SimOutDiss.g)
out.CoIs$SimOutFilt.g = ifelse(is.na(out.CoIs$SimOutFilt.g), 0.0, out.CoIs$SimOutFilt.g)
out.CoIs$SimOutSmeto.g = out.CoIs$SimOutDiss.g + out.CoIs$SimOutFilt.g

# Cumulative OUT
out.CoIs$CumOutDiss.g = cumsum(out.CoIs$SimOutDiss.g)
out.CoIs$CumOutFilt.g = cumsum(out.CoIs$SimOutFilt.g)
out.CoIs$CumOutSmeto.g = out.CoIs$CumOutDiss.g + out.CoIs$CumOutFilt.g

# Balance
out.CoIs$BalMassDisch.g = out.CoIs$CumAppMass.g - out.CoIs$CumOutSmeto.g

# Mass fraction
massOUT = tail(out.CoIs$CumOutSmeto.g, n=1)
out.CoIs$FracMassOut = (out.CoIs$SimOutSmeto.g / massOUT)
out.CoIs$FracDeltaOut = (out.CoIs$SimOutSmeto.g / massOUT)*out.CoIs$diss.d13C
out.CoIs$FracDeltaOut = ifelse(is.na(out.CoIs$FracDeltaOut), 0.0, out.CoIs$FracDeltaOut)

BulkDeltaOut = sum(out.CoIs$FracDeltaOut)

```

The total mass discharged (up to Week 15) and bulk isotope signature (up to week 11) was:

```{r}
# Cummulative S-metolachlor [g] discharged
massOUT

# Bulk isotope signature
BulkDeltaOut
```




6. Testing a regression tree (ommitted for now)

```{r, echo=F, message=FALSE}

# Regression Tree Example 1

library("party")

#outCoIs.model <- out.CoIs[, c("chExtreme", "changeflux", "AveDischarge.m3.h", "Volume.m3", "TotMassOut.mg")]

#fit <- ctree(TotMassOut.mg ~ chExtreme + changeflux + AveDischarge.m3.h + Volume.m3, data=na.omit(outCoIs.model))
#plot(fit)

# The model does not improve with more variables. 

# Regression Tree Example 2

library("rpart")

# grow tree 
#fit <- rpart(TotMassOut.mg ~ AveDischarge.m3.h + Volume.m3 + chExtreme + changeflux,  method="anova", data=outCoIs.model)

#printcp(fit) # display the results 
#plotcp(fit) # visualize cross-validation results 
#summary(fit) # detailed summary of splits

# create additional plots 
#par(mfrow=c(1,2)) # two plots on one page 
#rsq.rpart(fit) # visualize cross-validation results  	

## plot tree 
#plot(fit, uniform=TRUE, main="Regression Tree Total Mass Discharged ")
#text(fit, use.n=TRUE, all=TRUE, cex=.8)

```



## Save files

```{r, message=FALSE}

head(out.CoIs)

write.csv2(out.CoIs, 
           'Data/WeeklyHydroContam_R.csv', row.names = F)

# out.CoIs = read.csv2("Data/WeeklyHydroContam_R.csv")
# out.CoIs$ti = as.POSIXct(out.CoIs$ti, "%Y-%m-%d %H:%M", tz = "EST")

```

