---
title: "Rayleigh Calculations"
author: "PAZ"
date: "06/04/2017"
output: pdf_document
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
Sys.setlocale("LC_ALL", "English")
```

# Introduction

Degradation extent is calculated based on closed and open system Rayleigh equations

# Packages

```{r, warning=FALSE,}
library(sm)
library(vioplot)

library(dplyr)
library(tidyr)
library(zoo)
library(reshape)
library(ggplot2)
library("ggrepel")

library("plotly")
library("cowplot")
library("gridExtra")
library("Cairo")
library("GGally")
library("scales")

library("plotKML")

# Stats
library("vegan")
library("cluster")

# Saving a xlxs file
# library(xlsx)
```

# Lab parameters

```{r}
# Initial signature measured in tank
initialDelta = d13Co = -32.253

# Define initial concentration (for Raleigh plots)
#Co <- 8 # ug/g dry soil (based on Corn applications)
# Co <- 6.53 # ug/g dry soil (based on Max conc. measured in soils)
# Note: Each transect now has individual starting concentration

rho = 0.98*10^6 # soil density [g/m3]
depth = 0.01 # [m]

# Reference values
# Ehssan values:
epsilon_mean= -2.2 # ± 0.4
dE <- 0.4
epsilon_max = epsilon_mean + dE 
epsilon_min = epsilon_mean - dE


epsilon_lab = epsilon_mean

# Calculated in this Book from Bulk signatures and bulk concentrations
epsilonField_max = -1.3 + 0.24 
epsilonField_min = -1.3 - 0.24  
epsilonField_mean = -1.3 # ± 0.24

```

# Soils

```{r}

VERTICAL = F
if (VERTICAL) {
  # Mixed transects in one column (NO bulk)
  soils = read.csv2("Data/WeeklySoils_Rng.csv", 
                         na.strings=c('#DIV/0!', '', 'NA'), header = TRUE)
  soils$Date.ti <- as.POSIXct(strptime(soils$Date.ti, 
                                            "%Y-%m-%d %H:%M", tz="EST")) # csv typos, option 1
  sum(is.na(soils$Date.ti)) == 0
  
  names(soils)
  
  keepSoils <- c(
    "ID","Transect","Wnum",  "Date.ti", 
    "Conc.mug.g.dry.soil", "Conc.ComSoil.SD" ,
    "comp.d13C", "comp.d13C.SD", "comp.IMP.d13C", 
    "DD13C.comp"
  )
} else {
  
  soils = read.csv2("Data/MassBalance_R.csv", # File contains both soils & waters
                       na.strings=c('#DIV/0!', '', 'NA'), header = TRUE)
  names(soils)
  soils$ti <- as.POSIXct(strptime(soils$ti, "%Y-%m-%d %H:%M", tz="EST"))
  colnames(soils)[colnames(soils) == "ti"] <- "Date.ti"
  
  keepSoils <- c(
    "Date.ti", "WeekSubWeek", "ID.N",
    "Area.N", "Area.T", "Area.S",
    "iniCo.ug.g.N", "iniCo.ug.g.T", "iniCo.ug.g.S",
    "CumAppMass.g.N", "CumAppMass.g.T", "CumAppMass.g.S",
    "diss.d13C", "SD.d13C", 
    "CumAppMass.g" , 
    "CatchMassSoil.g", "CatchMassSoil.g.SD", 
    "comp.d13C.North" , "comp.d13C.SD.North", 
    "comp.d13C.Talweg", "comp.d13C.SD.Talweg", 
    "comp.d13C.South" , "comp.d13C.SD.South",
    "BulkCatch.d13", "BulkCatch.d13.SD",
    "Conc.mug.g.dry.soil.N", "Conc.mug.g.dry.soil.T", "Conc.mug.g.dry.soil.S",  
    "BulkCatch.Conc", "iniCo.Bulk"
  )
}
# Test
sum(is.na(soils$Date.ti)) == 0

soils <- soils[, colnames(soils) %in% keepSoils]
```


## Rayleigh (closed system, Elsner's notation)


$$ ln (\frac{1000 + \delta ^{13}C_0 + \Delta\delta^{13}C }{1000 + \delta^{13} C_0 }) = (\alpha - 1) \cdot ln f = \frac{\epsilon}{1000} \cdot ln f $$
were, 

$$ f = \frac{C_t}{C_0} $$

### Accounting for dilution

The Rayleigh equation above assumes that $f$ reflects solely reduction in concentrations due to degradation and should thus be expressed as $f_{degradation}$. Accounting for dilution processes, the remaining fraction that is measured in the field sample becomes then $f_{total}$, where:

$$ f_{total} = f_{degradation} \cdot f_{dilution} $$

Following Van Breukelen [-@VanBreukelen2007],

$$f_{degradation} = f_{total} \cdot F$$

where the dilution factor $F$ (i.e. the number of times the source volume has become diluted at the observation location) can be calculated if $\epsilon_{lab}$ is known:

$$F = e^{(\Delta/\epsilon_{lab}-lnf_{total})}$$

were,

$$\Delta = 1000 \cdot ln \Big( \frac{ 10^{-3} \delta^{13}_t C + 1}{10^{-3} \delta^{13}_0 C + 1}  \Big)$$
# Soils

### Initial soil concentration(s)

Estimation of initial concentrations followed breaking down the fraction of the catchment receiving application by transect, then 

```{r}

if (VERTICAL) {
  iniCo <- 12.17
# soils$iniCo <- 12.17 # Average for plots with application 1+ day
# soils$iniCo <- 9.96 # Median for plots with application 1+ day
# soils$iniCo <- 22.7 # MAX for plots with application 1+ day

# Based on mean plot Day +1, by Transect
# soils$iniCo <- ifelse(soils$Transect == "N", 16.2, 
#                          ifelse(soils$Transect == "T", 8.9, 
#                                 ifelse(soils$Transect == "S", 10.0, NA)))

} else {
# Already defined in book 05  
}


```

### Dilution factor ($F$)

```{r}

if (VERTICAL) {
  soils$ftot <- soils$Conc.mug.g.dry.soil/soils$iniCo

  # Van Breukelen notation
  soils$Delta <- 1000*log( (10^-3*soils$comp.d13C+1)/(10^-3*d13Co+1) )
  
  # Elsner notation
  # soils$DeltaX <- log(1000+d13Co+soils$DD13C.comp)/(1000+d13Co)
  
  soils$Fdil = 
    exp( soils$Delta/epsilon_lab -log(soils$ftot) ) 
  median(soils$Fdil, na.rm = T)
  
  # Fdil < 1, otherwise this 
  soils$Fdil <- ifelse(soils$Fdil < 1, NA, soils$Fdil)
  
  # In case lab E is not appropriate, an adjustment may be necessary
  adj_epsilon = -3.5
  
  #soils$FdilAdj = 
  #  exp( soils$Delta/adj_epsilon -log(soils$ftot) ) 
} else {
  soils$ftot.N <-soils$Conc.mug.g.dry.soil.N/soils$iniCo.ug.g.N
  soils$ftot.T <-soils$Conc.mug.g.dry.soil.T/soils$iniCo.ug.g.T
  soils$ftot.S <-soils$Conc.mug.g.dry.soil.S/soils$iniCo.ug.g.S
  
  soils$ftot.Bulk <-soils$BulkCatch.Conc/soils$iniCo.Bulk
  
  # Van Breukelen notation
  soils$Delta.N <- 1000*log( (10^-3*soils$comp.d13C.North +1)/(10^-3*d13Co+1) )
  soils$Delta.T <- 1000*log( (10^-3*soils$comp.d13C.Talweg +1)/(10^-3*d13Co+1) )
  soils$Delta.S <- 1000*log( (10^-3*soils$comp.d13C.South +1)/(10^-3*d13Co+1) )
  soils$Delta.Bulk <- 1000*log( (10^-3*soils$BulkCatch.d13 +1)/(10^-3*d13Co+1) )
  
  soils$Fdil.N = 
    exp( soils$Delta.N/epsilon_lab -log(soils$ftot.N) ) 
  soils$Fdil.T = 
    exp( soils$Delta.T/epsilon_lab -log(soils$ftot.T) ) 
  soils$Fdil.S = 
    exp( soils$Delta.S/epsilon_lab -log(soils$ftot.S) ) 
  
  soils$Fdil.Bulk = 
        exp( soils$Delta.Bulk/epsilon_lab -log(soils$ftot.Bulk) ) 
  median(soils$Fdil.N, na.rm = T)
  median(soils$Fdil.T, na.rm = T)
  median(soils$Fdil.S, na.rm = T)
  median(soils$Fdil.Bulk, na.rm = T)
  
  # soils$Fdil <- ifelse(soils$Fdil < 1, NA, soils$Fdil)
}

```

We can now obtain $f_{dilution}$ and $f_{degradation}$:

```{r}

if (VERTICAL){
  soils$fdil <- 1/soils$Fdil
  soils$fdeg <- soils$ftot * soils$Fdil 
  
  #soils$fdilAdj <- 1/soils$FdilAdj
  #soils$fdegAdj <- soils$ftot * soils$FdilAdj 
  
  soils$Dprct <- (1- soils$fdil)*100
  #soils$DprctAdj <- (1- soils$fdilAdj)*100
  
  soils$Bprct <- (1-soils$fdeg)*100
  #soils$BprctAdj <- (1-soils$fdegAdj)*100
  
  soils$Tprct <- (1-soils$ftot)*100
} else {
  soils$fdil.N <- 1/soils$Fdil.N
  soils$fdeg.N <- soils$ftot.N * soils$Fdil.N 
  
  soils$Dprct.N <- (1- soils$fdil.N)*100
  soils$Bprct.N <- (1-soils$fdeg.N)*100
  soils$Tprct.N <- (1- (1 - soils$Bprct.N/100) * (1 - soils$Dprct.N/100)  )*100
  
  # Talweg
  soils$fdil.T <- 1/soils$Fdil.T
  soils$fdeg.T <- soils$ftot.T * soils$Fdil.T 
  
  soils$Dprct.T <- (1- soils$fdil.T)*100
  soils$Bprct.T <- (1-soils$fdeg.T)*100
  soils$Tprct.T <- (1-soils$ftot.T)*100

  #South
  soils$fdil.S <- 1/soils$Fdil.S
  soils$fdeg.S <- soils$ftot.S * soils$Fdil.S 
  
  soils$Dprct.S <- (1- soils$fdil.S)*100
  soils$Bprct.S <- (1-soils$fdeg.S)*100
  soils$Tprct.S <- (1-soils$ftot.S)*100
  
  # Bulk
  soils$fdil.Bulk <- 1/soils$Fdil.Bulk
  soils$fdeg.Bulk <- soils$ftot.Bulk * soils$Fdil.Bulk 
  
  soils$Dprct.Bulk <- (1- soils$fdil.Bulk)*100
  soils$Bprct.Bulk <- (1-soils$fdeg.Bulk)*100
  soils$Tprct.Bulk <- (1-soils$ftot.Bulk)*100
  
}


```

Calculating degradation and respective statistical variation errors in B, according to [@Thulnner2012]: 

$$ | \Delta B | = \frac{1}{| \epsilon |} \cdot (100 - B[\%]) \cdot \sqrt{(ln(1-B[\%]/100))^2 \cdot (\Delta \epsilon)^2 + (\Delta \delta_x)^2 + (\Delta \delta_0)^2} $$

```{r}
# Deg effective
soils$DD.N <-  soils$comp.d13C.North - d13Co
soils$DD.T <-  soils$comp.d13C.Talweg - d13Co
soils$DD.S <-  soils$comp.d13C.South - d13Co
soils$DD.Bulk <- soils$BulkCatch.d13 - d13Co

soils$Belsner.N <- (1- ((1000 + d13Co + soils$DD.N)/(1000+ d13Co))^(1000/epsilon_lab) )*100
soils$Berr.Stat.N <- (1/-epsilon_lab)*
  (100-soils$Belsner.N)*( (log(1-soils$Belsner.N/100))^2 * dE^2 + (soils$comp.d13C.SD.North)^2 + (.5)^2 )^(1/2)

soils$Belsner.T <- (1- ((1000 + d13Co + soils$DD.T)/(1000+ d13Co))^(1000/epsilon_lab) )*100
soils$Berr.Stat.T <- (1/-epsilon_lab)*
  (100-soils$Belsner.T)*( (log(1-soils$Belsner.T/100))^2 * dE^2 + (soils$comp.d13C.SD.Talweg)^2 + (.5)^2 )^(1/2)

soils$Belsner.S <- (1- ((1000 + d13Co + soils$DD.S)/(1000+ d13Co))^(1000/epsilon_lab) )*100
soils$Berr.Stat.S <- (1/-epsilon_lab)*
  (100-soils$Belsner.S)*( (log(1-soils$Belsner.S/100))^2 * dE^2 + (soils$comp.d13C.SD.South)^2 + (.5)^2 )^(1/2)

soils$Belsner.Bulk <- (1- ((1000 + d13Co + soils$DD.Bulk)/(1000+ d13Co))^(1000/epsilon_lab) )*100
```


Calculating errors from systematic variation in input parameters, in this case $\epsilon_{lab}$:

```{r}
soils$Belsner.BulkField <- (1- ((1000 + d13Co + soils$DD.Bulk)/(1000+ d13Co))^(1000/epsilon_field) )*100
soils$Bdiff.LabField <- soils$Belsner.BulkField - soils$Belsner.Bulk
```


```{r}
soils$Berr.Stat.Bulk <- (1/-epsilon_lab)*
  (100-soils$Belsner.Bulk)*( (log(1-soils$Belsner.Bulk/100))^2 * dE^2 + (0.5)^2 + (.5)^2 )^(1/2)

soils$Berr.Stat.Field <- (1/-epsilon_field)*
  (100-soils$Belsner.BulkField)*( (log(1-soils$Belsner.BulkField/100))^2 * dE^2 + (0.5)^2 + (.5)^2 )^(1/2)
```



As well as the breakdown ($B*$), dilution factors ($D*$):

$$B^* = \frac{log(f_{dil})}{log(f_{tot})} $$

and their relationship $D*/B*$ which relates the extent of dilution relative to degradation.

```{r}
if (VERTICAL){
  
  
  #soils$DstarAdj = log(soils$fdilAdj)/log(soils$ftot)
  #soils$BstarAdj = log(soils$fdegAdj)/log(soils$ftot)
} else {
  soils$Dstar.N = log(soils$fdil.N)/log(soils$ftot.N)
  soils$Bstar.N = log(soils$fdeg.N)/log(soils$ftot.N)
  soils$DB.N = soils$Dstar.N/soils$Bstar.N
  
  soils$Dstar.T = log(soils$fdil.T)/log(soils$ftot.T)
  soils$Bstar.T = log(soils$fdeg.T)/log(soils$ftot.T)
  soils$DB.T = soils$Dstar.T/soils$Bstar.T
  
  soils$Dstar.S = log(soils$fdil.S)/log(soils$ftot.S)
  soils$Bstar.S = log(soils$fdeg.S)/log(soils$ftot.S)
  soils$DB.S = soils$Dstar.S/soils$Bstar.S

  soils$Dstar.Bulk = log(soils$fdil.Bulk)/log(soils$ftot.Bulk)
  soils$Bstar.Bulk = log(soils$fdeg.Bulk)/log(soils$ftot.Bulk)
  soils$DB.Bulk = soils$Dstar.Bulk/soils$Bstar.Bulk

  median(soils$DB.N, na.rm = T)
  median(soils$DB.T, na.rm = T)
  median(soils$DB.S, na.rm = T)

  mean(soils$DB.N, na.rm = T)
  mean(soils$DB.T, na.rm = T)
  mean(soils$DB.S, na.rm = T)

    
  ggplot(data = soils , aes(x=Date.ti))+
    geom_point(aes(y = Dstar.N, colour = "D*N")) + 
    geom_point(aes(y = Dstar.T, colour = "D*T")) +
    geom_point(aes(y = Dstar.S, colour = "D*S")) +
    theme_bw() 
  

  ggplot(data = soils , aes(x=Date.ti))+
    geom_point(aes(y = Bstar.N, colour = "B*N")) + 
    geom_point(aes(y = Bstar.T, colour = "B*T")) +
    geom_point(aes(y = Bstar.S, colour = "B*S")) +
    theme_bw()

}


  
```

Calculating the cummulative mass that was degraded and transpoted off-site can be done by converting the change in concentration to the change in mass, with the following conversion:

$$ \Delta M_t = M_t - M_0 $$
don't think this is correct! Should be B% instead of B*...

$$  M_{deg(t)} = \Delta M \cdot B^* $$
(thinking about dilution in the wrong way? ->  Dilution factor can only be applied to concentrations, not masses) 

because,

$$B^* = 1 - D^*$$

$$ M_t ~ [\mu g] = C_t ~ [\mu g/ g~soil] * \rho ~ [g~soil / m^3 ] * A_{T} [m^2] * D~[m]    $$
Note here that $A_T$ refers to the representative catchment area (subcatchment) of the given transect, and depth is 0.01 $m$. 

```{r}

soils$dM.g.N =  (soils$iniCo.ug.g.N*rho*soils$Area.N*depth - soils$Conc.mug.g.dry.soil.N*rho*soils$Area.N*depth)*1/10^6
soils$dM.g.T =  (soils$iniCo.ug.g.T*rho*soils$Area.T*depth - soils$Conc.mug.g.dry.soil.T*rho*soils$Area.T*depth)*1/10^6
soils$dM.g.S =  (soils$iniCo.ug.g.S*rho*soils$Area.S*depth - soils$Conc.mug.g.dry.soil.S*rho*soils$Area.S*depth)*1/10^6

# Doubt, Don't think this formula is applicable to mass calculations, 
# as mass is not affected by dilution, only concentration is
CONC = FALSE
if (CONC) {
  soils$Mdeg.g.N <- soils$dM.g.N*soils$Bstar.N
  soils$Mdeg.g.T <- soils$dM.g.T*soils$Bstar.T
  soils$Mdeg.g.S <- soils$dM.g.S*soils$Bstar.S
  soils$SumDeg.g <- soils$Mdeg.g.N + soils$Mdeg.g.T + soils$Mdeg.g.S

  soils$Mdil.g.N <- soils$dM.g.N*soils$Dstar.N
  soils$Mdil.g.T <- soils$dM.g.T*soils$Dstar.T
  soils$Mdil.g.S <- soils$dM.g.S*soils$Dstar.S
  soils$SumDil.g <- soils$Mdil.g.N + soils$Mdil.g.T + soils$Mdil.g.S
  
} else {
  soils$Mdeg.g.N <- soils$CumAppMass.g.N*soils$Belsner.N/100
  soils$Mdeg.g.T <- soils$CumAppMass.g.T*soils$Belsner.T/100
  soils$Mdeg.g.S <- (soils$CumAppMass.g.S)*soils$Belsner.S/100
  soils$SumDeg.g <- soils$Mdeg.g.N + soils$Mdeg.g.T + soils$Mdeg.g.S
}

soils$Mrem.g <- 
  (soils$Conc.mug.g.dry.soil.N*rho*soils$Area.N*depth +
     soils$Conc.mug.g.dry.soil.T*rho*soils$Area.T*depth +
     soils$Conc.mug.g.dry.soil.S*rho*soils$Area.S*depth)*1/10^6

soils$Mrem.g.N <- soils$Conc.mug.g.dry.soil.N*rho*soils$Area.N*depth*1/10^6
soils$Mrem.g.T <- soils$Conc.mug.g.dry.soil.N*rho*soils$Area.T*depth*1/10^6
soils$Mrem.g.S <- soils$Conc.mug.g.dry.soil.N*rho*soils$Area.S*depth*1/10^6

soils$Mloss.g.N <- soils$CumAppMass.g.N - (soils$Mdeg.g.N + soils$Mrem.g.N)
soils$Mloss.g.T <- soils$CumAppMass.g.T - (soils$Mdeg.g.T + soils$Mrem.g.T)
soils$Mloss.g.S <- soils$CumAppMass.g.S - (soils$Mdeg.g.S + soils$Mrem.g.S)

soils$PrctRem.N <- (soils$Mrem.g.N/soils$CumAppMass.g.N)*100
soils$PrctRem.T <- (soils$Mrem.g.T/soils$CumAppMass.g.T)*100
soils$PrctRem.S <- (soils$Mrem.g.S/(soils$CumAppMass.g.S))*100

soils$PrctDeg <- (soils$SumDeg.g/soils$CumAppMass.g)*100 # Sum of transect degradations relative to Applied mass
soils$PrctDeg.N <- (soils$Mdeg.g.N/soils$CumAppMass.g.N)*100
soils$PrctDeg.T <- (soils$Mdeg.g.T/soils$CumAppMass.g.T)*100
soils$PrctDeg.S <- (soils$Mdeg.g.S/(soils$CumAppMass.g.S))*100

soils$PrctRem <- (soils$Mrem.g/soils$CumAppMass.g)*100
soils$PrctLoss <- 100 - (soils$PrctRem + soils$PrctDeg)
soils$PrctLoss.N <- 100 - (soils$PrctRem.N + soils$PrctDeg.N)
soils$PrctLoss.T <- 100 - (soils$PrctRem.T + soils$PrctDeg.T)
soils$PrctLoss.S <- 100 - (soils$PrctRem.S + soils$PrctDeg.S)

names(soils)
dropS <- c("Area.Catch", "Area.N", "Area.T", "Area.S", 
            "WeekSubWeek", 
           # Isotopes Soils
           "comp.d13C.North", "comp.d13C.SD.North",
           "comp.d13C.Talweg", "comp.d13C.SD.Talweg"  ,
           "comp.d13C.South", "comp.d13C.SD.South",
           "BulkCatch.d13", "BulkCatch.d13.SD",
           "DD.N", "DD.T", "DD.S","DD.Bulk", 
           # ISotopes Water
           # "diss.d13C", "SD.d13C",
           # Concentrations
           "BulkCatch.Conc" ,
           "Conc.mug.g.dry.soil.N", "Conc.mug.g.dry.soil.T", "Conc.mug.g.dry.soil.S", 
      
           "iniCo.Bulk", "iniCo.ug.g.N" , "iniCo.ug.g.T",  "iniCo.ug.g.S",
           
           # Mass
           "CatchMassSoil.g", "CatchMassSoil.g.SD",
           
           # VanBreukelen
           "Fdil.N", "Fdil.T", "Fdil.S", "Fdil.Bulk",
            "ftot.N","ftot.T", "ftot.S", "ftot.Bulk",  
           "fdil.N", "fdil.T", "fdil.S", "fdil.Bulk",
           "fdeg.N", "fdeg.T", "fdeg.S", "fdeg.Bulk",
           "Bstar.N", "Bstar.T", "Bstar.S", "Bstar.Bulk", 
           "Dstar.N", "Dstar.T", "Dstar.S", "Dstar.Bulk", 
           "dM.g.N", "dM.g.T", "dM.g.S"# ,
           #"Mdeg.g.N", "Mdeg.g.T", "Mdeg.g.S"
           )

soilsRay <- soils[ , !colnames(soils) %in% dropS]
soilsRay <- soilsRay[complete.cases(soilsRay[ , "ID.N"]), ]
soilsRay$ID.N.x <- soilsRay$ID.N
names(soilsRay)

mean(soilsRay$Bdiff.LabField[3:8])
mean(soilsRay$Bdiff.LabField[12:13])
mean(soilsRay$Berr.Stat.Bulk[3:8])
mean(soilsRay$Berr.Stat.Bulk[12:13])

mean(soils$Berr.Stat.Field, na.rm = T)

names(soilsRay)
```

Calculating a field enrichment after correcting for via breakdown factor ($B^*$),

$$\epsilon_{field} = B^* \cdot \epsilon_{lab} = \frac{\Delta}{ln f _{total} } $$

```{r}
soils$Efield.N <- soils$Bstar.N * epsilon_lab
soils$Efield.T <- soils$Bstar.T * epsilon_lab
soils$Efield.S <- soils$Bstar.S * epsilon_lab
soils$Efield.Bulk <- soils$Bstar.Bulk * epsilon_lab
# soils$EfieldAdj <- soils$BstarAdj * epsilon_lab

Efield.Bulk <- mean(soils$Efield.Bulk, na.rm = T)
Efield.Bulk

sd(soils$Efield, na.rm = T)

# median(soils$EfieldAdj)
```


## Save soils and water for Bar Plots

```{r}
# Check correct:
soils$Etrue =  soils$Efield.Bulk/soils$Bstar.Bulk
median(soils$Etrue, na.rm = T) # Should be close to -2.2, YES.
soils$Etrue <- NULL
#write.csv2(soils, 
#           'Data/Rayleigh_Soils.csv', row.names = F)

write.csv2(soilsRay, 
           'Data/Rayleigh_Soils.csv', row.names = F)



```


In Van Breuklen, the degraded and diluted fractions are plotted against each other. They find that in the fringe of the plume, where more dilution occurs, also more degradation occurs, likely associated to oxidant availability and lower toxicity levels.   

```{r}

if (VERTICAL){
  # Van Breuklen plot this 
  soils$degY = -log(soils$fdeg)
  soils$dilX = -log(soils$fdil)
  
  
  DBmodel<-lm( degY ~ dilX , data= soils, subset=(!is.na(Etrue) )) 
  cof_DB <- as.numeric(coef(DBmodel)[2])
  # se_DB <- summary(DBmodel)$coef[[4]]*1000
  summary(DBmodel)
  
  ggplot(data = subset(soils, ( !is.na(fdil) & dilX>0 )), aes(x=dilX, y=degY)) +
    geom_point()

}

```


In contrast, in top soils, a slightly negative but not significant relationship between extent of dilution and degradation was found. This is to be expected as concentration in tops soils are lower than in aquifer systems for legacy contaminants. At lower concentrations, lack of sufficient substrate may be associated to lower bacterial communities capable of degradation. 


# Waters

Conversion of initial concentration in soils to pore water, assuming all S-met mass is available, may lead to an Fdil_w. 

For waters, no dilution factor can be applied. The degraded fraction of off-site transport must be equivalent to catchment soils, as degradation is negligible within an event. However, it may be of interest to obtain the degradation extent that would be computed if only outlet observations were conducted, without knowledge of dilution extent in catchment soils. As such, the fraction degraded in outlet waters will be obtained via $\epsilon_{lab}$ and with the closed system Rayleigh equation, without making use of concentration data. 


```{r}
waters = read.csv2("Data/WeeklyHydroContam_R.csv")
waters$ti <- as.POSIXct(strptime(waters$ti, "%Y-%m-%d %H:%M", tz="EST"))
colnames(waters)[colnames(waters) == "ti"] <- "Date.ti"
waters$Events <- factor(waters$Events, levels = unique(waters$Events))
waters$Event <- factor(waters$Event, levels = unique(waters$Event))

names(waters)

keepWater <- c(
  "Date.ti", "WeekSubWeek", "Sampled" , "Volume.m3", "AveDischarge.m3.h",
  "Markers" , "TimeDiff", "Duration.Hrs", 
  "Conc.mug.L", "Conc.SD" ,
  "OXA_mean", "OXA_SD", 
  "ESA_mean", "ESA_SD", 
  "diss.d13C", "SD.d13C",
  "filt.d13C", "filt.SD.d13C",
  "DD13C.diss", "DD13C.filt",  
  "Appl.Mass.g",  "CumAppMass.g",
  "DissSmeto.g", "DissSmeto.g.SD", 
  "DissOXA.g", "DissOXA.g.SD",
  "DissESA.g", "DissESA.g.SD",
  "FiltSmeto.g", "FiltSmeto.g.SD",
  "TotSMout.g",  "TotSMout.g.SD", 
  "MELsm.g", "MELsm.g.SD",
  "CumOutDiss.g", "CumOutFilt.g",
  "CumOutSmeto.g", "CumOutMELsm.g", 
  "Events"
)
waters <- waters[ , colnames(waters) %in% keepWater]
```

## Field Assumptions

Converting soil to volumetric concentrations in soils, 

$$ C_{soil}~[\mu g/L_{soil}] = C_{soil}[\frac{\mu g~_{S-met}}{g_{soil}}] \cdot\rho_{soil}[\frac{g_{soil}}{m^3}] \cdot \frac{ 1 m^3}{ 10^3 L_{soil}} $$


$$C_{soil}~[\mu g/L_{H_2O} ] = \frac{ C_{soil} }{\theta_{sat} + \rho_{soil} \cdot K_d } $$

```{r}
# S-metolachlor Mass [g]
# Conc. [ug/g dry soil] * [g/10^6 ug] * density [g/m3] * depth [m]* A [m2] 
# Soil bulk density: 2200 or 0.99? -> Leaching experiments: 0.98 [g/cm3]
rho = 0.98*10^6 # soil density [g/m3]
depth = 0.01 # [m]
theta = 0.4
# Kd = 2.397/10^6 # m3/g (aged)
Kd = 3.99/10^6 #m3/g (fresh)
```



```{r}

# Cosed vs. Open system Rayleigh
OPEN = T

if (OPEN) {
  waters$f.diss <- 
  ((10^(-3)*waters$diss.d13C + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_lab))
  
  waters$f.diss.min <- 
    ((10^(-3)*waters$diss.d13C + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_max))
  
  waters$f.diss.max <- 
    ((10^(-3)*waters$diss.d13C + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_min))
  
  waters$B.diss <- (1-waters$f.diss)*100
  waters$B.diss.max <- (1-waters$f.diss.min)*100
  waters$B.diss.min <- (1-waters$f.diss.max)*100
} else if (!VERTICAL){
  ws <- merge(waters, soils, by = "Date.ti", all = T )
  
  # Assume Bulk soils conc. doesn't evolve close to event
  ws$BulkCatch.Conc <- na.locf(ws$BulkCatch.Conc)
  
  # Convert concentrations from mass to vol H20, assuming linear sorption
  ws$poolCo_w <- (ws$BulkCatch.Conc*rho/10^3)/(theta + rho*Kd)
  ws$iniCo_w <- (iniCo*rho/10^3)/(theta + rho*Kd)
  
  # f_tot
  # Problem here is initial Co is not applied but, available at time of discharge
  ws$ftot_w <- ws$Conc.mug.L/ws$poolCo_w

  # Van Breukelen notation
  ##########################
  ##########################
  # Do we need to change d13Co to initial at event or initial product ??
  ws$Delta_w <- 1000*log( (10^-3*ws$diss.d13C.x + 1)/(10^-3*d13Co+1) )
  
  ws$Fdil_w = 
    exp( ws$Delta_w/epsilon_lab -log(ws$ftot_w) ) 
  median(ws$Fdil_w, na.rm = T)
  
  # Fdil < 1, otherwise this 
  ws$Fdil_w <- ifelse(ws$Fdil_w < 1, NA, ws$Fdil_w)
  
  ws$fdil_w <- 1/ws$Fdil_w
  ws$fdeg_w <- ws$ftot_w * ws$Fdil_w 
  
  ws$Dprct_w <- (1- ws$fdil_w)*100
  #ws$DprctAdj <- (1- ws$fdilAdj)*100
  
  ws$Bprct_w <- (1-ws$fdeg_w)*100
  #ws$BprctAdj <- (1-ws$fdegAdj)*100
  
  ws$Tprct_w <- (1-ws$ftot_w)*100
  
  ws$Dstar_w = log(ws$fdil_w)/log(ws$ftot_w)
  ws$Bstar_w = log(ws$fdeg_w)/log(ws$ftot_w)
  
  ws$Dstar_w  <- ifelse(ws$Dstar_w< 0, NA, ws$Dstar_w)
  ws$Bstar_W  <- ifelse(ws$Bstar_w> 1, NA, ws$Bstar_w)
  
  ws$DB_w = ws$Dstar_w/ws$Bstar_w
}



```

## Save Waters for Bar Plots

```{r}
write.csv2(waters, 
           'Data/Rayleigh_Waters.csv', row.names = F)
```

