---
title: "PNAS Figures"
author: "PAZ"
date: "22 novembre 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
Sys.setlocale("LC_ALL", "English")
```

## Required R-packages:

```{r, message=FALSE}

# Data wrangling
library("plyr")
library("dplyr")

# Melting data sets & changin axes
library("reshape2")
library("ggrepel")

# Plotting:
library("ggplot2")
library("cowplot")
library("gridExtra")
library("Cairo")
library("GGally")
library("scales")

```

## Working directory

```{r, message=FALSE}

# setwd("D:/Documents/these_pablo/Alteckendorf2016/R")
# setwd("/Users/DayTightChunks/Documents/PhD/Routput/Alteck/R")
# setwd("D:/Documents/these_pablo/Alteckendorf2016/00_TransparencyFolder")
getwd()

# Show all test graphs (change to TRUE)
SHOW = FALSE
CHECK_ERR = FALSE
```

## Lab & field parameters
```{r}

# Pure and cuve isotope average
initialDelta = d13Co = -32.235

# Lab enrichment:
# Alteck
# More degr. under saturation because anaerobic is best according to Fatima
epsilon_max = -1.5 # +/- 0.3 (@ 20C, 20% vwc) 
epsilon_min = -2.0 # +/- 0.2 (@ 20C, 40% vwc) 

epsilon_mean = -1.75 # Lab
sd(c(epsilon_max, epsilon_min))

field_epsilon = -1.692547 # Field
field_epsilon_sd = 0.363319 # ±

```


## Soils

```{r}
weeklySoil = read.csv2("Data/WeeklySoils_Rng.csv", na.strings=c('#DIV/0!', '', 'NA'), header = TRUE)
weeklySoil$Date.ti <- as.POSIXct(strptime(weeklySoil$Date.ti, "%Y-%m-%d %H:%M", tz="EST")) # csv typos, option 1
#weeklySoil$Date.ti <- as.POSIXct(strptime(weeklySoil$Date.ti, "%d/%m/%Y %H:%M", tz="EST"))
sum(is.na(weeklySoil$Date.ti))

#weeklySoil$Conc.ComSoil.SD <- 
#  ifelse(weeklySoil$Conc.ComSoil.SD == as.character("#DIV/0!"), NA, as.numeric(as.character(weeklySoil$Conc.ComSoil.SD)))
                                     
str(weeklySoil)

# weeklySoil = weeklySoil %>%
#  group_by(Transect) %>%
#  arrange(Transect, Wnum)

weeklySoil$Transect <- factor(weeklySoil$Transect, levels = c("N", "T", "S"))


```

## Soil Concentrations

```{r}
#, fig.height=3, fig.width=3}

###############################
# Concentrations
###############################
###############################
###############################
#weeklySoil$ti[3] <- as.POSIXct("2016-04-14 08:25:00")
#weeklySoil$ti[14] <- as.POSIXct("2016-04-14 08:25:00")
#weeklySoil$ti[24] <- as.POSIXct("2016-04-14 08:25:00")
#lb1a2 <- paste("App.")
lbW012 <- paste("App.W0/1/2")
lbW9 <- paste("App.W9")

limits_conc_soil <- aes(ymin=Conc.mug.g.dry.soil-Conc.ComSoil.SD,  ymax=Conc.mug.g.dry.soil+Conc.ComSoil.SD)
#limits_conc_soil <- aes(ymin=mean-0.5, ymax=mean+0.5)

pd <- position_dodge(0.5) # move them .05 to the left and right

co = ggplot(weeklySoil[4:48, ], 
           aes(x=Date.ti, y=Conc.mug.g.dry.soil, colour=Transect, group = Transect)) + 
  
  geom_point() +
  geom_line() +
  
  # Error bars
  geom_errorbar(limits_conc_soil, width=.1, position=pd) +
  # scale_y_continuous(limits=c(0,10),oob = rescale_none) +
  
  # Themes and axes
  theme_bw() +
  theme(legend.position = "top",
        axis.text.x=element_text(angle = 45, hjust = 1)
        #axis.text.x=element_blank(), 
        #axis.title.x=element_blank()
        ) +
  
  # scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  ylab(expression(paste("Conc. S-Meto.  ", {({mu}*g / g.soil.dry)}))) +
  # facet_wrap(~Transect, nrow = 3) +
  # xlab("Date") +
  # theme() +
  scale_x_datetime(breaks = date_breaks("1 weeks"), labels = date_format("%b %d")) +
  
  # Smooth linear models
  # stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  # stat_smooth(method = "lm") +
  
  # Text
  # W0 Application
  # annotate("text", x = as.POSIXct('2016-03-25 08:04:00'), y = 4, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), y = 0.5, xend = as.POSIXct('2016-03-26 01:04:00'), yend = -0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +

  # W1 Application
  geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), y = 0.5, 
                   xend = as.POSIXct('2016-04-05 08:04:00'), yend = 0), color = "black", 
               arrow = arrow(length = unit(0.2, "cm"))) +
  # W2 Application
  annotate("text", x = as.POSIXct('2016-04-15 08:04:00'), y = 1, label = lbW012, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), y = 0.5, 
                   xend = as.POSIXct('2016-04-13 08:04:00'), yend = 0), color = "black", 
               arrow = arrow(length = unit(0.2, "cm"))) +
  # W9 Application
  annotate("text", x = as.POSIXct('2016-05-26 08:04:00'), y = 4.5, label = lbW9, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-05-26 08:04:00'), y = 4, 
                   xend = as.POSIXct('2016-05-25 18:04:00'), yend = 0), color = "black", 
               arrow = arrow(length = unit(0.2, "cm"))) +

  geom_text_repel(aes(label=as.factor(Wnum)),
                 size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                 max.iter = 2e3,
                nudge_x = .05) 
co

# Linear model
# ggsave(co, filename = "CompositeConcLM.png", width = 7, height = 5, units = "in", scale = 1)

## ggsave(co, filename = "CompositeConcLM.tiff", height = 10, width = 8.7, units = 'cm')

# No linear model
# ggsave(co, filename = "CompositeConc.png", width = 7, height = 5, units = "in", scale = 1)

```

# Soil isotope signatures

```{r, fig.height=6, fig.width=6}

initialDelta
weeklySoil$DD13C.comp <- (weeklySoil$comp.d13C - (initialDelta))

#weeklySoil$ngC.Label <- ifelse(weeklySoil$ngC.mean<5, "< 5 ng", 
#                               ifelse(weeklySoil$ngC.mean<10, "< 10 ng", 
#                               ifelse((weeklySoil$ngC.mean >= 10 & weeklySoil$ngC.mean < 15), "< 15 ng", 
#                               ifelse((weeklySoil$ngC.mean >= 15 & weeklySoil$ngC.mean < 20), "< 20 ng",
#                               ifelse(weeklySoil$ngC.mean >= 20 & weeklySoil$ngC.mean < 30, "< 30 ng C", "> 30 ng")))))

limits_dCsoil <- aes(ymin=comp.d13C-comp.d13C.SD, ymax=comp.d13C+comp.d13C.SD)
#limits_dCsoil <- aes(ymin=comp.d13C-0.5, ymax=comp.d13C+0.5)
lb1a <- paste("App.~S-meto.")
lb1ab <- paste("delta^{13}~C:-32.25")
lb1a2 <- paste("App. ")

lbW012 <- paste("App.W0/1/2")
lbW9 <- paste("App.W9")

if (SHOW) {
ggplot(weeklySoil, aes(x=Date.ti, y=comp.d13C, colour=Transect, group = Transect)) + 
  geom_errorbar(limits_dCsoil, width=.05) +
  geom_point() +
  theme_bw() +
  #stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  stat_smooth(method = "lm") +
  facet_wrap(~Transect, nrow = 3) +
  xlab("Date") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1)) +
  #ylab(expression(paste({delta}^"13","C", ' \211'))) +
  ylab(expression(paste({delta}^"13","C", ' (\u2030)'))) +
  scale_y_continuous(breaks=seq(-34,-21,2)) +
  geom_hline(yintercept = -31.21, color = "dodgerblue4", linetype = "dotted") +
  geom_hline(yintercept = -30.71, color = "dodgerblue3", linetype = "dotted") +
  geom_hline(yintercept = -31.71, color = "dodgerblue3", linetype = "dotted") +
  annotate("text", x = as.POSIXct('2016-04-05 22:04:00'), y = -22.5, label = lb1a, parse = T, size = 3.0) +
  annotate("text", x = as.POSIXct('2016-04-05 22:04:00'), y = -23.5, label = lb1ab, parse = T, size = 3.0) +
  
  annotate("text", x = as.POSIXct('2016-03-25 08:04:00'), y = -29, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-03-25 08:04:00'), y = -29.8, 
                   xend = as.POSIXct('2016-03-25 08:04:00'), yend = -31.0), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("text", x = as.POSIXct('2016-04-03 00:04:00'), y = -29, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-03 00:04:00'), y = -29.8, 
                   xend = as.POSIXct('2016-04-05 08:04:00'), yend = -31.0), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("text", x = as.POSIXct('2016-04-13 08:04:00'), y = -25, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-13 08:04:00'), y = -26, 
                   xend = as.POSIXct('2016-04-13 08:04:00'), yend = -31.0), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("text", x = as.POSIXct('2016-05-26 08:04:00'), y = -29, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-05-26 08:04:00'), y = -29.8, 
                   xend = as.POSIXct('2016-05-25 08:04:00'), yend = -31.0), 
               arrow = arrow(length = unit(0.2, "cm"))) +
  #scale_x_continuous(breaks=seq(0,11,1)) +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  
  #annotate("text", x = as.POSIXct('2016-05-30 20:04:00'), y = -30.5, label = lb1a, parse = T, size = 2.0) +
  theme(legend.position = "top")
  
# Linear model (LM)
# ggsave(isCo, filename = "CompositeIsotopesLM.png", width = 7, height = 5, units = "in", scale = 1)
# No linear model
# ggsave(isCo, filename = "CompositeIsotopes.png", width = 7, height = 5, units = "in", scale = 1)

}

if (SHOW) {
# View(weeklySoil)
# Ommitted, graph is tautological.
### Delta vs. f (Soils)
ggplot(weeklySoil, aes(x=f.comp, y=DD13C.comp, colour=Transect, group = Transect)) + 
  #geom_errorbar(limits_dCsoil, width=.05) +
  geom_point() +
  theme_bw() +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  #stat_smooth(method = "lm") +
  facet_wrap(~Transect, nrow = 3) +
  scale_x_reverse() +
  xlab("Fraction remaining (f)") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1)) +
  #ylab(expression(paste({delta}^"13","C", ' \211'))) +
  ylab(expression(paste({Delta~delta}^"13","C", ' (\u2030)'))) +
  #scale_y_continuous(breaks=seq(-34,-21,2)) +
  theme(legend.position = "top") +
  #geom_text_repel(aes(label=WeekNo, color = factor(Transect)), 
                  #arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  #force = 1, 
                  #point.padding = unit(1.0, 'lines'), 
                  #max.iter = 2e3,
                  #nudge_x = .2) +
  geom_point() 
  
  
  ###############################
###############################
### DeltaDelta vs time
###############################
# View(weeklySoil)
# limits_DdCsoil <- aes(ymin=comp.d13C-comp.d13C.SD-initialDelta, ymax=comp.d13C+comp.d13C.SD-initialDelta)
limits_DdCsoil <- aes(ymin=comp.d13C-comp.d13C.SE-initialDelta, ymax=comp.d13C+comp.d13C.SE-initialDelta)

# pd <- position_dodge(0.5)
# AOdf[1:27,]
deltaTime = ggplot(na.omit(weeklySoil), aes(Date.ti, DD13C.comp)) +
  geom_errorbar(limits_DdCsoil) +
  geom_point(aes(shape = Transect)) +
  labs(shape="Transect", colour = "Mass Carbon") +
  
  # geom_point(weeklySoil[1:48, ], aes(x=Date.ti, y=DD13C.comp, colour=Transect, group = Transect)) +
  
  # Themes and axes
  theme_bw() +
  theme(# legend.position="none",
        # axis.title.x = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +

  
  ylab(expression(paste({Delta~delta}^"13","C", ' (\u2030)'))) +
  scale_y_continuous(breaks=seq(0, 8, 1)) +
  # ylab(expression(paste({delta}^"13","C", ' \211'))) +
  # ylab(expression(paste({delta}^"13","C", ' (\u2030)'))) +
  # facet_wrap(~Transect, nrow = 3) +
  
  # Smooth linear models
  stat_smooth(data = weeklySoil, aes(group = Transect), method = "lm", se=FALSE) +
  # stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  # stat_smooth(data=subset(weeklySoil[4:27, ]), method = "lm", formula = y~x, se=F) +
  # stat_smooth(data=subset(weeklySoil[18:36, ]), method = "lm", formula = y~x, se=F) +
  
  # Text
  # Application W0
  annotate("text", 
           x = as.POSIXct('2016-04-04 01:04:00'), y = 6, label = lbW012, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-03 08:04:00'), y = 5.5, 
                   xend = as.POSIXct('2016-03-25 22:04:00'), yend = -0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +
  #annotate("text", 
  #         x = as.POSIXct('2016-04-03 00:04:00'), y = 2, label = lb1a2, parse = T, size = 3.0) +
  
  # Application W1
  geom_segment(aes(x = as.POSIXct('2016-04-03 08:04:00'), y = 5.5, 
                   xend = as.POSIXct('2016-04-05 08:04:00'), yend = 0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +
  # annotate("text", x = as.POSIXct('2016-04-15 08:04:00'), y = 1, label = lb1a2, parse = T, size = 3.0) +
  
  # Application W2
  geom_segment(aes(x = as.POSIXct('2016-04-03 08:04:00'), y = 5.5, 
                   xend = as.POSIXct('2016-04-13 08:04:00'), yend = 0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +
  
  # Application W9
  annotate("text", 
           x = as.POSIXct('2016-06-10 08:04:00'), y = 1.8, label = lbW9, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-06-10 08:04:00'), y = 1.5, 
                   xend = as.POSIXct('2016-05-25 18:04:00'), yend = 0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +
  
  geom_text_repel(aes(label=Wnum), 
                 arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                 force = 1, 
                 point.padding = unit(1.0, 'lines'), 
                 max.iter = 2e3,
                 nudge_x = .2)

deltaTime

}

```

## Delta time final

```{r}
# , fig.height=3, fig.width=3}

limits_DdCsoil <- aes(ymin=comp.d13C-comp.d13C.SD-initialDelta, ymax=comp.d13C+comp.d13C.SD-initialDelta, colour = Transect)

# weeklySoil$ngC.Label <- ifelse(weeklySoil$ngC.mean < 10, "< 10 ng", "> 10 ng")
                               
                               # ifelse( weeklySoil$ngC.mean >= 10 & weeklySoil$ngC.mean < 15, "< 15 ng", "> 15 ng"))
wk <- weeklySoil
# wk <- na.omit(weeklySoil)
wk$Application <- ifelse(wk$Date.ti == as.POSIXct('2016-05-17 09:16:00', tz = "EST") & wk$Transect == "T", 0.05, 
                    ifelse(wk$Date.ti == as.POSIXct('2016-03-30 12:18:00', tz = "EST") & wk$Transect == "T", 0.05, 
                     ifelse(wk$Date.ti == as.POSIXct('2016-04-05 15:08:00', tz = "EST") & wk$Transect == "T", 0.05,
                       ifelse(wk$Date.ti == as.POSIXct('2016-04-14 13:52:00', tz = "EST") & wk$Transect == "T", 0.05, NA))))

deltaTime = ggplot(wk, aes(Date.ti, DD13C.comp)) +
  geom_errorbar(limits_DdCsoil, width = 0.5) +
  geom_point(aes(colour = Transect), size = 2) +
  labs(colour="Transect")+ #, shape = "Mass Carbon") +
  geom_point(aes(Date.ti, Application, shape = "Application"), 
             size =2 , data = wk,  legend.title = element_blank()) +
  # Themes and axes
  theme_bw() +
  theme(legend.position="top",
        # axis.title.x = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  guides(col = guide_legend(nrow =  3), shape = guide_legend(nrow =  3)) +
  ylab(expression(paste({Delta~delta}^"13","C", ' (\u2030)'))) +
  scale_y_continuous(breaks=seq(0, 8, 1)) +
  stat_smooth(data = weeklySoil, method = "lm", se=T) +
  geom_text_repel(data = subset(weeklySoil), aes(label=Wnum, colour = Transect),
                 arrow = arrow(length = unit(0.01, 'npc'), type = "closed"),
                 force = 1, 
                 point.padding = unit(0.5, 'lines'), 
                 max.iter = 2e3,
                 nudge_x = .5, show.legend = F)

deltaTime

#ggsave(deltaTime, filename = "Composite_DD.png", width = 8, height = 5, units = "in", scale = 1)

```


## Degradation

```{r, fig.height=6, fig.width=6}

lb1a2 <- paste("App.")

lb1b <- paste("(A)~epsilon:-1.5")
lb1b2 <- paste("(B)~epsilon:-2.0")

if (SHOW) {
  ggplot(weeklySoil, aes(x=Date.ti, y=B.min.comp, colour=Transect, group = Transect)) +
    ylab("Degradation (%)") +
    scale_y_continuous(breaks=seq(0, 100, 20)) +
    geom_point() +
    # geom_point(aes(x=Date.ti, y=B.comp, colour=Transect, group = Transect)) +
    theme_bw() +
    stat_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) + 
    annotate("text", x = as.POSIXct('2016-04-11 20:04:00'), 
             y = 100, label = lb1b, parse = T, size = 3.0, color = "grey40") +
    annotate("text", x = as.POSIXct('2016-04-11 20:04:00'), 
             y = 85, label = lb1b2, parse = T, size = 3.0, color = "dodgerblue4" ) +
    
    annotate("text", x = as.POSIXct('2016-03-25 08:04:00'), y = 30, label = lb1a2, parse = T, size = 3.0) +
    geom_segment(aes(x = as.POSIXct('2016-03-25 08:04:00'), y = 25, 
                     xend = as.POSIXct('2016-03-25 08:04:00'), yend = 20), 
                 arrow = arrow(length = unit(0.2, "cm"))) +
    annotate("text", x = as.POSIXct('2016-04-03 00:04:00'), y = 30, label = lb1a2, parse = T, size = 3.0) +
    geom_segment(aes(x = as.POSIXct('2016-04-03 00:04:00'), y = 25, 
                     xend = as.POSIXct('2016-04-05 08:04:00'), yend = 20), 
                 arrow = arrow(length = unit(0.2, "cm"))) +
    annotate("text", x = as.POSIXct('2016-04-13 08:04:00'), y = 30, label = lb1a2, parse = T, size = 3.0) +
    geom_segment(aes(x = as.POSIXct('2016-04-13 08:04:00'), y = 25, 
                     xend = as.POSIXct('2016-04-13 08:04:00'), yend = 20), 
                 arrow = arrow(length = unit(0.2, "cm"))) +
    annotate("text", x = as.POSIXct('2016-05-26 08:04:00'), y = 30, label = lb1a2, parse = T, size = 3.0) +
    geom_segment(aes(x = as.POSIXct('2016-05-26 08:04:00'), y = 25, 
                     xend = as.POSIXct('2016-05-25 08:04:00'), yend = 20), 
                 arrow = arrow(length = unit(0.2, "cm"))) +
    scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
    #scale_x_continuous(breaks=seq(0,11,1)) +
    theme(legend.position = "top") 
}

```


## Water 

```{r}
AOdf = read.csv2("Data/WeeklyHydroContam_R.csv")
str(AOdf)

AOdf$ti <- as.POSIXct(strptime(AOdf$ti, "%Y-%m-%d %H:%M", tz="EST"))
sum(is.na(AOdf$ti)) == 0

```

##  Outlet - Concentrations

```{r} 
#, fig.height=6, fig.width=6}
# Volumes sampled vs. not sampled
vols <- AOdf %>%
  group_by(Sampled) %>%
  summarise_each(funs(sum(., na.rm=TRUE)), Volume.m3)

prctSampled <- vols[2, "Volume.m3"]/(vols[1, "Volume.m3"] + vols[2, "Volume.m3"])*100
prctSampled

prctNotSampled <- vols[1, "Volume.m3"]/(vols[1, "Volume.m3"] + vols[2, "Volume.m3"])*100
prctNotSampled

# Subset the data
# newdata <- mydata[ which(mydata$gender=='F' & mydata$age > 65), ]

limits_conc <- aes(ymin=Conc.mug.L-Conc.SD, ymax=Conc.mug.L+Conc.SD, color = Event, group = Event)
dfSampled <- AOdf[which(AOdf$Sampled == 'Sampled'), ]

dfSampled$Row <- seq.int(nrow(dfSampled))

conc1 <- ggplot(dfSampled, aes(x=ti, y=Conc.mug.L)) +
  geom_point( aes(color = Event, group = Event)) +
  # geom_point( aes(color = Event)) +
  
  # Error bars
  geom_errorbar(limits_conc, width=1) +
  
  # Themes and axes
  theme_bw() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1), 
        #axis.text.x=element_blank(),
        #axis.title.x=element_blank(),
        legend.position="top"
        )+
  guides(col = guide_legend(nrow =  2)) +  # Sets legend parameters
  xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  ylab(expression(paste("Conc. S-Meto.  ", {({mu}*g / L)}))) +
  scale_y_continuous( breaks = c(28, 26, 18,16,14,12,10, 8, 6, 4, 2, 0), limits = c(0, 30) ) +
  
  # Smooth linear models
  geom_smooth(data=subset(AOdf[4:28, ]), method = "lm", formula = y ~ poly(x, 2), se = F) +
  geom_smooth(data=subset(AOdf[28:length(AOdf), ]), method = "lm", formula = y ~ poly(x, 2), se= F) +
  #stat_smooth(data=subset(dfSampled[19:30, ]), method = "lm", formula = y ~ x) +
  # Text
  # Application W9
  # annotate("text", 
  #         x = as.POSIXct('2016-06-10 08:04:00'), y = -1, label = lbW9, parse = T, size = 3.0) +
  # geom_segment(aes(x = as.POSIXct('2016-06-05 08:04:00'), y = -1, 
  #                 xend = as.POSIXct('2016-05-25 18:04:00'), yend = -0.9), color = "black",
  #             arrow = arrow(length = unit(0.2, "cm"))) +
  geom_text_repel(aes(label=Events, color = Event), # WeekSubWeek or Weeks
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05, show.legend = F)
conc1

cor.p.SM <- cor.test(dfSampled$Conc.mug.L, dfSampled$diss.d13C, method = "pearson")
cor.p.SM
cor.p.ESA <- cor.test(dfSampled$ESA_mean, dfSampled$diss.d13C)
cor.p.ESA

cor.p.SM.soils <- cor.test(wk$Conc.mug.g.dry.soil, wk$comp.d13C)
cor.p.SM.soils
```


##  Outlet - Masses vs time

```{r}
library("tidyr")

dfSampled1 <- dfSampled[ , c("ti", "Peak", "Event", "Events",
                   "TotSMout.g", "TotSMout.g.SD", 
                   "DissOXA.g", "DissOXA.g.SD", 
                   "DissESA.g", "DissESA.g.SD",
                   "MELsm.g", "MELsm.g.SD")]

```


## Outlet Isotope Shifts (DD)

In the same plot consider this secondary axis, where the secondary axis is a formulat of the first:

ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  scale_y_continuous(
    "mpg (US)", 
    sec.axis = sec_axis(~ . * 1.20, name = "mpg (UK)")
  )

Or this: https://github.com/tidyverse/ggplot2/wiki/Align-two-plots-on-a-page

```{r, fig.height=11, fig.width=7}

if (SHOW) {
  limits_conc <- aes(ymin=Conc.mug.L-Conc.SD, ymax=Conc.mug.L+Conc.SD, color = Event, group = Event)
  ggplot(AOdf[28:length(AOdf),], aes(x=ti, y=Conc.mug.L)) +
    geom_point( aes(color = Weeks, group = Weeks)) +
    # Error bars
    # geom_errorbar(aes(ymin=mean.d13C-SD.d13C, ymax=mean.d13C+SD.d13C), width=.1) +
    # geom_errorbar(limits_conc, width=1) +
    
    # Themes & axes
    # theme_gray() +
    theme_bw() +
    theme(legend.position = "none") +
    theme(axis.text.x=element_text(angle = 45, hjust = 1),
          axis.text.y = element_blank(),
          legend.title = element_blank(),
          plot.margin = unit(c(0,3.5,0,0), "lines")) +
    #scale_x_datetime(breaks = date_breaks("week"), labels = date_format("%m/%d")) +
    scale_y_continuous(breaks = c(20,15,10,5,0), limits = c(-5, 20) ) +
    xlab("Date") +
    ylab("") +
    
    # Smooth linear models
      stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
    #geom_hline(yintercept = -31.21, color = "dodgerblue4", linetype = "dotted") +
    #geom_hline(yintercept = -30.71, color = "dodgerblue3", linetype = "dotted") +
    #geom_hline(yintercept = -31.71, color = "dodgerblue3", linetype = "dotted") +
    
    # Text
    #annotate("text", x = as.POSIXct('2016-06-25 00:04:00'), y = -31.2, label = lb1, parse = T) +
    annotate("text", x = as.POSIXct('2016-05-27 08:04:00'), y = -3, label = "App.4", parse = T) +
    geom_segment(aes(x = as.POSIXct('2016-05-26 08:04:00'), y = -4, 
                     xend = as.POSIXct('2016-05-26 08:04:00'), yend = -5.0), 
                 arrow = arrow(length = unit(0.2, "cm"))) +
    geom_text_repel(aes(label=Weeks, color = factor(Weeks)),
                    size = 3,
                    arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                    force = 0.5, 
                    point.padding = unit(0.5, 'lines'), 
                    max.iter = 2e3,
                    nudge_x = .05)
 
}

```


## Outlet Isotopes - Continous (by Week)

```{r}
AOdf$SD.d13C.err  <- ifelse(is.na(AOdf$SD.d13C), 0.5, AOdf$SD.d13C)
# limits_dC <- aes(ymin=diss.d13C-SD.d13C.err, ymax=diss.d13C+SD.d13C.err, color = Weeks, group = Weeks)
limits_dC <- aes(ymin=diss.d13C-SD.d13C, ymax=diss.d13C+SD.d13C, color = Weeks, group = Weeks)
# View(AOdf)

iso <- ggplot(AOdf, aes(x=ti, y=diss.d13C)) +
  #geom_errorbar(aes(ymin=mean.d13C-SD.d13C, ymax=mean.d13C+SD.d13C), width=.1) +
  geom_errorbar(limits_dC, width=1) +
  #theme_gray() +
  theme_bw() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1)) +
  #scale_x_datetime(breaks = date_breaks("week"), labels = date_format("%m/%d")) +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  geom_point( aes(color = Weeks, group = Weeks)) +
  #stat_smooth(method = "lm", formula = y ~ x) +
  geom_smooth(data=subset(AOdf[4:length(AOdf), ]), method = "lm", formula = y ~ poly(x, 2)) +
  #stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  #theme(axis.text.x = element_blank()) +
  #theme(plot.margin = unit(c(1,1,1,1), "lines")) +
  geom_hline(yintercept = -31.21, color = "dodgerblue4", linetype = "dotted") +
  geom_hline(yintercept = -30.71, color = "dodgerblue3", linetype = "dotted") +
  geom_hline(yintercept = -31.71, color = "dodgerblue3", linetype = "dotted") +
  #annotate("text", x = as.POSIXct('2016-06-25 00:04:00'), y = -31.2, label = lb1, parse = T) +
  xlab("Date") +
  #theme(legend.position="top") +
  scale_y_continuous(breaks = c(-32,-31,-30,-29, -28, -27), limits = c(-32, -26.4) ) +
  ylab(expression(paste({delta}^"13","C", ' (\u2030)'))) +
  geom_text_repel(aes(label=WeekSubWeek, color = factor(Weeks)),
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05, show.legend = F) +
  geom_text_repel(aes(label=Events),
                  size = 3,
                  arrow = arrow(length = unit(0.01, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(1.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .06, show.legend = F)
  #ylab(expression(paste({delta}^"13","C", ' \211')))
  #ylab(expression(paste({delta}^"13","C"))) 
 
iso

# ggsave(iso, filename = "Outlet_Delta_ti_cont.png", width = 8, height = 5, units = "in", scale = 1)

```


## All plots

```{r,  fig.height=7.5, fig.width=7.01}

if (CHECK_ERR) {
concSoils <- co + theme(legend.position='none')
concWater <- conc1 + theme(legend.position='none')
isoSoils <- deltaTime + theme(legend.position='none')

legend_soils = get_legend(deltaTime)
legend_water = get_legend(conc1)

grid4 <- plot_grid(concSoils, isoSoils, concWater, iso2, 
                    ncol =2, nrow = 2, align ="v",
                    labels = c("A", "C", "B", "D"))
fig1 <- ggdraw() +
  draw_plot(grid4, x=0, y=.2, width =  1, height = .8) + 
  draw_plot(legend_water, x=0.5, y = 0.05, width = 0.5, height = 0.2) +
  draw_plot(legend_soils, x=0, y=0.05, width = 0.6, height = 0.2)

fig1
}

#ggsave(fig1, filename = "SoilsAndOutlet.tiff", height = 18, width = 17.8, units = 'cm')

```

## Mass balance approach

```{r}
library("ggplot2")
library("scales")
library("reshape2")
library("zoo")

soilsOut = read.csv2("Data/MassBalance_R.csv", header = T)
soilsOut$ti <- as.POSIXct(soilsOut$ti, "%Y-%m-%d %H:%M", tz = "EST")
sum(is.na(soilsOut$ti))

# Remove bulk catchment values that came from 
# at least one source of inputed data (all inputed data has no stdrd. devs)
soilsOut$BulkCatch.d13 <- ifelse(is.na(soilsOut$comp.d13C.SD.North), NA, 
                                 ifelse(is.na(soilsOut$comp.d13C.SD.Talweg), NA,
                                              ifelse(is.na(soilsOut$comp.d13C.SD.South), NA, soilsOut$BulkCatch.d13)
                                        )
                                 )


print("Mass Balance Soils")
str(soilsOut)

names(soilsOut)
# Melt data set
##Subset the necessary columns
soilsRemainMass <- soilsOut[, c("ti" , "WeekSubWeek", 
                                "diss.d13C", "SD.d13C", 
                                "CumAppMass.g", "CatchMassSoil.g",
                                "CumOutDiss.g", "CumOutFilt.g", "CumOutMELsm.g"
                                # "f.mean.bulk", "B.mean.bulk"
                                )]

# Fraction remaining with mean lab epsilon (Alteck)
soilsRemainMass$f.mean.outlet <- 
  ((10^(-3)*soilsRemainMass$diss.d13C + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_mean))

soilsRemainMass$B.mean.outlet <-
  (1 - soilsRemainMass$f.mean.outlet)*100

```

### As bar chart, showing only summary when soils are measured

```{r}
# Delete rows with only cumulative obs.
# n specifies max no. of NA's in a row allowed
delete.na <- function(DF, n=0) { 
  DF[rowSums(is.na(DF)) <= n , ] 
}

remainOnSampleDay <- delete.na(soilsRemainMass, n = 5)

# Omit rows whre no soil samples were made 
# This implies missing data values for discharge
remainOnSampleDay <- subset(remainOnSampleDay, !is.na(CatchMassSoil.g)  ) #| !is.na(B.mean.outlet)


remainOnSampleDay$Persist.Prct <- (remainOnSampleDay$CatchMassSoil.g/remainOnSampleDay$CumAppMass.g)*100
remainOnSampleDay$TPs.PrctOut <- ((remainOnSampleDay$CumOutMELsm.g-
                                  (remainOnSampleDay$CumOutDiss.g+remainOnSampleDay$CumOutFilt.g))
                                  /remainOnSampleDay$CumAppMass.g)*100
remainOnSampleDay$SM.PrctOut <- ((remainOnSampleDay$CumOutDiss.g+remainOnSampleDay$CumOutFilt.g)
                                 /remainOnSampleDay$CumAppMass.g)*100

remainOnSampleDay$Unknown <- 100 - 
  (remainOnSampleDay$Persist.Prct + 
     remainOnSampleDay$TPs.PrctOut + 
     remainOnSampleDay$SM.PrctOut)

remainOnSampleDay$F.Bulk <- remainOnSampleDay$f.mean.bulk*100
remainOnSampleDay$B.Bulk <- remainOnSampleDay$B.mean.bulk

remainOnSampleDay$F.Outlet <- remainOnSampleDay$f.mean.outlet*100
remainOnSampleDay$B.Outlet <- remainOnSampleDay$B.mean.outlet



keepPrct <- c("ti",
            "Persist.Prct", "TPs.PrctOut", "SM.PrctOut", "Unknown",
            "F.Bulk" #, "F.Outlet" # , "B.Bulk",  "B.Outlet"
            )

keepPrctCSIA <- c("ti",
            "Persist.Prct", 
            "SM.PrctOut",
            "B.Bulk", "F.Bulk" #, "F.Outlet" #,  "B.Outlet"
            )

prctMB <- remainOnSampleDay[ , (names(remainOnSampleDay) %in% keepPrct)]

# Computing CSIA-based Mass Balance
# Not that %s have been calculated from survey applied mass.
prctCSIA <- remainOnSampleDay[ , (names(remainOnSampleDay) %in% keepPrctCSIA)]
prctCSIA$Tot <- prctCSIA$Persist.Prct + prctCSIA$B.Bulk + prctCSIA$SM.PrctOut 
prctCSIA$Error <- ifelse((prctCSIA$Tot - 100) < 0 , 0, prctCSIA$Tot - 100)
prctCSIA$Persist.Prct <- ifelse(prctCSIA$Error > 0, 
                                prctCSIA$Persist.Prct - prctCSIA$Error , 
                                prctCSIA$Persist.Prct) 

# Chosen to substract error from the persistent fraction method
# as this method required larger assumptions:
# e.g. Extrapolation to catchment areas & GC-MS variability

prctCSIA$Unknown <- ifelse((100 - prctCSIA$Tot) < 0 , 0, 100 - prctCSIA$Tot)
prctCSIA$Tot <- NULL

if (prctMB$ti[3] == as.POSIXct("2016-03-28 22:38:00", tz = "EST")) {
  prctMB <- prctMB[4:nrow(prctMB), ]
}

prctMB <- melt(prctMB, id=c("ti"))
prctCSIA <- melt(prctCSIA, id=c("ti"))
levels(prctMB$variable)
prctMB$variable <- factor(prctMB$variable, levels = c( "SM.PrctOut", "TPs.PrctOut", "Unknown", "Persist.Prct",
                                                       "F.Bulk" #, "F.Outlet" ,
                                                       # "B.Bulk", "B.Outlet"
                                                       ))
levels(prctCSIA$variable)
prctCSIA$variable <- factor(prctCSIA$variable, levels = c( "Error", "SM.PrctOut" , "B.Bulk", "Unknown",  "Persist.Prct",
                                                           "F.Bulk"
                                                       ))

prctCSIA <- within(prctCSIA, value[variable == 'Error' & value == 0] <- NA)
 
#massbar <- ggplot(data = prctMB , aes(x=ti, y=value))+
massbar <- ggplot(data = prctMB , aes(x=ti, y=value))+
  theme_bw() +
  # geom_bar(stat = "identity", aes(fill = variable)) +
  #geom_bar(data = subset(prctMB, !(variable %in% c("F.Bulk","B.Bulk","F.Outlet", "B.Outlet"))), 
  #         aes(x=ti, y=value, fill = variable ),
  #         stat = "identity") +
  geom_bar(data = subset(prctMB, 
                         variable != "F.Bulk" # & variable != "B.Bulk" 
                         # & variable != "F.Outlet" # & variable != "B.Outlet"
                           ), 
           aes(x=ti, y=value, fill = variable ),
           stat = "identity") +
  geom_point(data = subset(prctMB, 
                           variable == "F.Bulk"  # |  variable == "B.Bulk"
                           # | variable == "F.Outlet" # | variable == "B.Outlet"
                           ), 
             aes(x=ti, y=value, shape = factor(variable) )) +
  stat_smooth(data=subset(prctMB, 
                           variable == "F.Bulk"), 
              method = "lm", formula = y ~ poly(x, 2), se = T, aes(col = 'F.Bulk') , alpha = 0.1, size=0.2) +
  # stat_smooth(data=subset(prctMB, 
  #                         variable == "F.Outlet"), 
  #            method = "lm", formula = y ~ poly(x, 2), se = T, aes(col = 'F.Outlet') , alpha = 0.1, size=0.2) +
  
  # Add error bars,
  # see: http://stackoverflow.com/questions/30872977/how-to-stack-error-bars-in-a-stacked-histogram-using-geom-errorbar-ggplot
  # geom_errorbar(aes(ymin=value-0.5, ymax=value+0.5), width=.5, position = "identity")+
  xlab("Date") +
  ylab("% Fraction SM") + 
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  theme(# legend.position="top"
        # axis.title.x = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  scale_fill_manual( 
                    #values = c("#a6611a", "#dfc27d", "#80cdc1", "#018571"),
                    values = c("#80cdc1", "#018571", "#a6611a", "#dfc27d"),
                    name="Mass Balance", # \n
                       breaks=c("SM.PrctOut", "TPs.PrctOut", "Unknown", "Persist.Prct"),
                       labels=c("SM (Outlet)", "TPmeq (Outlet)", "Unknown", "SM (Top soil)" )) +
  scale_shape(name="CSIA", labels = c("Non-degraded (Bulk)"# , "Non-degraded (Outlet)"
                                      )) +
  scale_color_manual( name= "CSIA (LM)", 
                      values = c("black"#, "dodgerblue"
                                 ), 
                      labels = c("Non-degraded (Bulk)" #, "Non-degraded (Outlet)"
                                 )
                      ) 


  #scale_fill_brewer(# palette="OrRd", 
  #                  palette = c("#a6611a", "chocolate", "green4", "dodgerblue"),
  #                  # values=c("#999999", "chocolate", "green4", "dodgerblue"), 
  #                  name="% Fraction", # \n
  #                     breaks=c("Unknown", "Persist.Prct",  "SM.PrctOut", "TPs.PrctOut"),
  #                     labels=c("Unknown", "SM (Top soil)", "SM (Outlet)", "TPs (Outlet)" ))
  
massbar
# ggsave(massbar, filename = "MassBalBar.png", width = 8, height = 5, units = "in", scale = 1)


# CSIA Based approach
csbar <- ggplot(data = prctCSIA , aes(x=ti, y=value))+
  theme_bw() +
  # geom_bar(stat = "identity", aes(fill = variable)) +
  #geom_bar(data = subset(prctMB, !(variable %in% c("F.Bulk","B.Bulk","F.Outlet", "B.Outlet"))), 
  #         aes(x=ti, y=value, fill = variable ),
  #         stat = "identity") +
  geom_bar(data = subset(prctCSIA, 
                         variable != "F.Bulk" & variable != "Error" 
                           ), 
           aes(x=ti, y=value, fill = variable ),
           stat = "identity") +
  geom_point(data = subset(prctCSIA, 
                         variable == "Error" 
                           ),
             aes(x=ti, y=value, shape = factor(variable) )
             ) +
  # Add error bars,
  # see: http://stackoverflow.com/questions/30872977/how-to-stack-error-bars-in-a-stacked-histogram-using-geom-errorbar-ggplot
  # geom_errorbar(aes(ymin=value-0.5, ymax=value+0.5), width=.5, position = "identity")+
  xlab("Date") +
  ylab("% Fraction SM") + 
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  theme(# legend.position="top"
        # axis.title.x = element_blank(),
        axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  scale_fill_manual( 
                    #values = c("#a6611a", "#dfc27d", "#80cdc1", "#018571"),
                    values = c("#80cdc1", "#018571", "#a6611a", "#dfc27d"),
                    name="CSIA + Mass Balance", # \n
                       breaks=c("SM.PrctOut", "B.Bulk", "Unknown", "Persist.Prct"),
                       labels=c("SM (Outlet)", "Degraded (Top Soil)", "Unknown", "SM (Top soil)" )) +
  scale_shape_manual(name="", 
              values = (4), 
              labels = c("Error")) 


  #scale_fill_brewer(# palette="OrRd", 
  #                  palette = c("#a6611a", "chocolate", "green4", "dodgerblue"),
  #                  # values=c("#999999", "chocolate", "green4", "dodgerblue"), 
  #                  name="% Fraction", # \n
  #                     breaks=c("Unknown", "Persist.Prct",  "SM.PrctOut", "TPs.PrctOut"),
  #                     labels=c("Unknown", "SM (Top soil)", "SM (Outlet)", "TPs (Outlet)" ))
  
csbar
# ggsave(csbar, filename = "MassBalCSIABar.png", width = 8, height = 5, units = "in", scale = 1)
```


### As continous graph

```{r, fig.height=5, fig.width=5}

# Rearrange data frame
remainMassMolten <- soilsRemainMass[, c("ti" ,"CumAppMass.g", "CumOutDiss.g", "CumOutFilt.g", "CumOutMELsm.g" , "CatchMassSoil.g")]
remainMassMolten <- melt(remainMassMolten, id=c("ti"))

pg <- remainMassMolten
pg <- na.omit(pg)

# Change variable names:
levels(pg$variable)[levels(pg$variable)=="CumAppMass.g"] <- "Applied SM Cum. (Survey)"
levels(pg$variable)[levels(pg$variable)=="CumOutMELsm.g"] <- "MEL-SM Cum. (Outlet)"
levels(pg$variable)[levels(pg$variable)=="CatchMassSoil.g"] <- "Persistent SM (Top soil 1cm)"

levels(pg$variable)[levels(pg$variable)=="CumOutDiss.g"] <- "Dissolved SM Cum. (Outlet)"
levels(pg$variable)[levels(pg$variable)=="CumOutFilt.g"] <- "Sediment SM Cum. (Outlet)"


# Change the order:
levels(pg$variable)
pg$variable <- factor(pg$variable, levels = c("Applied SM Cum. (Survey)",  "Persistent SM (Top soil 1cm)", "MEL-SM Cum. (Outlet)", "Dissolved SM Cum. (Outlet)", "Sediment SM Cum. (Outlet)" ))

pgSimple <- pg[which(pg$variable != ("Dissolved SM Cum. (Outlet)") & pg$variable != ("Sediment SM Cum. (Outlet)")), ]
# names(pg)[names(pg)=="variable"]  <- "Estimated Mass"

massBalTop <- ggplot(pg) + 
   geom_line(aes(x=ti, y=value, group = variable, color=variable)) +
  geom_point(aes(x=ti, y=value, group = variable, shape=variable, color=variable)) +

  # Themes and axes
  theme_bw() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1), 
        # axis.text.x=element_blank(),
        # axis.title.x=element_blank(),
        legend.position="top"
        )+
  # labs(group = "Estimated Mass") +
  guides(col = guide_legend(ncol = 1)) +  # Sets legend parameters
  scale_colour_manual(values=c("black" , "dodgerblue4", "violet", "chartreuse4" , "sandybrown")) +
  scale_shape_manual(values = c(8, 17, 25, 18, 7)) + 
  
  #scale_colour_manual(values=c("black" , "chartreuse4",  "orangered2",  "#F8766D", "dodgerblue4", "#619CFF", "#7CAE00")) +
  #scale_shape_manual(values = c(15, 18, 16,  23, 17, 13, 6)) + 
  
  xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  ylab(expression(paste("S-metolachlor Mass ", {(g)}))) +
  # scale_y_continuous(breaks = c(100, 5000, 10000, 20000), limits = c(100, 20000) ) 
  scale_y_continuous(trans=log_trans(), breaks=c(1,5,10,50,100,500,1000,5000, 10000))
massBalTop


massBalBottom <- ggplot(pg) + 
  geom_line(aes(x=ti, y=value, color=variable)) +
  
  # Themes and axes
  theme_bw() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1), 
        #axis.text.x=element_blank(),
        #axis.title.x=element_blank(),
        legend.position="none"
        )+
  # guides(col = guide_legend(nrows = 2)) +  # Sets legend parameters
  xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  ylab(expression(paste("Mass. S-Meto.  ", {(g)}))) +
  scale_y_continuous(breaks = c(1, 25, 50, 100), limits = c(0, 100) )


# massBal = plot_grid(massBalTop, massBalBottom, ncol = 1, nrow = 2, align = "v")



massBal_MEL <- ggplot(pgSimple) + 
  geom_line(aes(x=ti, y=value, group = variable, color=variable)) +
  
  # Themes and axes
  theme_bw() +
  theme(# axis.text.x=element_text(angle = 45, hjust = 1), 
        axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        legend.position="top"
        
        )+
  labs(color = "Estimated Mass") +
  guides(col = guide_legend(ncol = 3)) +  # Sets legend parameters
  
  # xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  ylab(expression(paste("S-metolachlor ", {(g)})))

# massBal_MEL

massBalLegend <- get_legend(massBalTop)

```


## Catchment degradation based on bulk signatures

```{r}

# Pure and cuve isotope average
d13Co

# Lab enrichment
epsilon_mean

if (SHOW) {
  # Remaining fraction
  soilsOut$DD13C.bulk <- (soilsOut$BulkCatch.d13 - (d13Co))
  
  # Max epsilon (30C, 20%) 
  soilsOut$f.max.bulk <-
    ((10^(-3)*soilsOut$BulkCatch.d13 + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_max))
  
  soilsOut$B.max.bulk <-
    (1 - soilsOut$f.max.bulk)*100 
  
  # Min epsilon (20C, 40%) 
  soilsOut$f.min.bulk <-
    ((10^(-3)*soilsOut$BulkCatch.d13 + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_min))
  
  soilsOut$B.min.bulk <-
    (1 - soilsOut$f.min.bulk)*100
  
  # Mean epsilon (# Alteck)
  soilsOut$f.mean.bulk <- 
    ((10^(-3)*soilsOut$BulkCatch.d13 + 1)/(10^(-3)*d13Co + 1))^(1000/(epsilon_mean))
  
  soilsOut$B.mean.bulk <-
    (1 - soilsOut$f.mean.bulk)*100
  
  
  
  ggplot(soilsOut, aes(x=ti, y=B.mean.bulk)) +
    geom_point() +
    # geom_point(aes(x=Date.ti, y=B.comp, colour=Transect, group = Transect)) +
    
    # Theme and axes
    theme_bw() +
    ylab("Degr. %") +
    theme(legend.position = "top",
          #axis.title = element_blank(),
          #axis.title.x = element_blank(),
          #axis.text.x = element_blank()
          axis.text.x=element_text(angle = 45, hjust = 1)
          ) +
    xlab("Date") +
    scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
    
    # scale_y_continuous(breaks = c(25, 50, 75, 100), limits = c(0, 100) ) +
    # stat_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) 
    # geom_smooth(data=subset(weeklySoil[14:28, ]), method = "lm", formula = y ~ poly(x, 2), se = F) +
    geom_smooth(aes(group = 1), method = "lm", formula = y ~ poly(x, 2))
    # stat_smooth(data=subset(weeklySoil[4:39, ]), method = "lm", formula = y ~ poly(x, 2), se = F) 
    # stat_smooth(method = "lm", formula = y ~ x, se=FALSE)
    #geom_text_repel(aes(label=Wnum, color = factor(Transect)),
     #               size = 3,
      #              arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
       #             force = 0.5, 
        #            point.padding = unit(0.5, 'lines'), 
         #           max.iter = 2e3,
          #          nudge_x = .05)
  
  # bulkB
}
```


# Degradation based on transect signatures 

```{r, fig.height=7.5, fig.width=7}

if (CHECK_ERR) {
# Merge Bulk Degradation into weekly soil dataframe (only one (e.g. South) transect as root)
soilsOut$PersistPrct <- (soilsOut$CatchMassSoil.g/soilsOut$CumAppMass.g)*100
soilsOut$DischPrct <- ((soilsOut$CumOutDiss.g+soilsOut$CumOutFilt.g)/soilsOut$CumAppMass.g)*100
bulkDegDF <-  soilsOut[, c("ti", "WeekSubWeek", "ID.S" ,"BulkCatch.d13", "B.mean.bulk", "B.max.bulk", "B.min.bulk", "PersistPrct", "DischPrct")]

bulkDegDF$TotBL <- bulkDegDF$B.min.bulk + bulkDegDF$PersistPrct + bulkDegDF$DischPrct
bulkDegDF$LeachPrct <- 100-bulkDegDF$TotBL
bulkDegDF$LeachPrctCorr <- ifelse(bulkDegDF$LeachPrct > 0, bulkDegDF$LeachPrct, NA)



# Delerte rows from specified Columns
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

bulkDegDF <- completeFun(bulkDegDF, "TotBL")

bulkDegDF$Transect <- "Degraded (Bulk)" 

# names(bulkDegDF)[names(bulkDegDF) == "B.mean.bulk"] <- "B.mean.com"
names(bulkDegDF)[names(bulkDegDF) == "ti"] <- "Date.ti" 

#Splitting the identifier name into Type, Week No., tc..
bulkDegDF$ID.S <- as.character(bulkDegDF$ID.S)
split <- strsplit(bulkDegDF$ID.S, "AW-S-", fixed = T)
bulkDegDF$Wnum <- sapply(split, "[", 2) # Creates new column without "Split0"

bulkDegDF$Week = paste("W", bulkDegDF$Wnum, sep = "")

bulkDegDF <- bulkDegDF[, c("Date.ti", "Transect", 
                           "B.mean.bulk",  "B.max.bulk",  "B.min.bulk", 
                           "PersistPrct", "DischPrct", "TotBL", "LeachPrct", "LeachPrctCorr",  
                           "Week")]
bulkDegDF$RemainLabel <- "Persistent frac. (Top Soil 1cm)"
bulkDegDF$DischLabel <- "Discharge (Outlet)"
bulkDegDF$LeachLabel <- "Leached (Inferred)"


levels(bulkDegDF$Transect)[levels(bulkDegDF$Transect)=="Bulk"] <- "Degraded (Bulk)"


wSoil <- weeklySoil[, c("Date.ti", "Transect", "B.mean.comp", "B.max.comp", "B.min.comp")]
levels(wSoil$Transect)[levels(wSoil$Transect)=="N"] <- "North"
levels(wSoil$Transect)[levels(wSoil$Transect)=="T"] <- "Talweg"
levels(wSoil$Transect)[levels(wSoil$Transect)=="S"] <- "South"

# colnames(wSoil) <- c("Date.ti", "Transect", "B.mean.bulk",  "B.max.bulk",  "B.min.bulk")

# wSoil$Week <- NA
# wSoilBulkDeg <- rbind(wSoil, bulkDegDF)

#names(bulkDegDF)[names(bulkDegDF) == "ID.S"] <- "ID"
#wSoilBulkDeg <- merge(weeklySoil, bulkDegDF, by="ID", all = T)
#wSoilBulkDeg$BulkLabel[!is.na(wSoilBulkDeg$WeekSubWeek)] <- "Bulk"


levels(wSoil$Transect)
#wSoil$Transect <- factor(wSoil$Transect, levels = c("Bulk Fraction", "North", "Talweg", "South" ))
wSoil$Transect <- factor(wSoil$Transect, levels = c("North", "Talweg", "South" ))

#wSoilBulkDeg$B.max.bulk[wSoilBulkDeg$Transect == "North"] <- NA
#wSoilBulkDeg$B.max.bulk[wSoilBulkDeg$Transect == "Talweg"] <- NA
#wSoilBulkDeg$B.max.bulk[wSoilBulkDeg$Transect == "South"] <- NA

#wSoilBulkDeg$B.min.bulk[wSoilBulkDeg$Transect == "North"] <- NA
#wSoilBulkDeg$B.min.bulk[wSoilBulkDeg$Transect == "Talweg"] <- NA
#wSoilBulkDeg$B.min.bulk[wSoilBulkDeg$Transect == "South"] <- NA

limits_bulkdeg <- aes(ymin=B.min.bulk, ymax=B.max.bulk, x = Date.ti, colour = Transect, group = Transect)

Bsoil1 <- 
  ggplot() +
  # geom_point(size = 2) +
  # ggplot(data = wSoilBulkDeg, aes(Date.ti, B.mean.bulk, colour = Transect, shape=Transect, group = Transect)) +
  geom_point(data = wSoil, 
             aes(Date.ti, B.mean.comp, colour = Transect, shape=Transect, group = Transect), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, B.mean.bulk, colour = Transect, shape=Transect, group = Transect), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, PersistPrct, colour = RemainLabel, shape=RemainLabel, group = RemainLabel), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, DischPrct, colour = DischLabel, shape=DischLabel, group = DischLabel), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, LeachPrctCorr, colour = LeachLabel, shape=LeachLabel, group = LeachLabel), size = 2) +
  
  geom_errorbar(data = bulkDegDF, limits_bulkdeg) + # With 2 data frames
  stat_smooth(data=subset(bulkDegDF, Transect == "Degraded (Bulk)"), 
              mapping = aes(y = B.mean.bulk, x = Date.ti), # With 2 data frames
              colour = " black",
              method = "lm", formula = y ~ poly(x, 2), se=F) +
  
  stat_smooth(data=subset(bulkDegDF, RemainLabel == "Persistent frac. (Top Soil 1cm)"), 
              mapping = aes(y = PersistPrct, x = Date.ti), # With 2 data frames
              colour = "darkblue",
              method = "lm", formula = y ~ poly(x, 2), se=F) +
  scale_colour_manual(values=c("black" , "chartreuse4",  "orangered2",  "#F8766D", "dodgerblue4", "#619CFF", "#7CAE00")) +
  scale_shape_manual(values = c(15, 18, 16,  23, 17, 13, 6)) + 
  theme_bw() +
  ylab("S-metolachlor Sinks (%)") +
  theme(legend.position = "top",
        legend.title = element_blank(),
        #axis.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        #axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  # xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  # ylab(expression(paste("Degradation %"))) +
  
  # scale_y_continuous(breaks = c(25, 50, 75, 100), limits = c(0, 100) ) +
  
  # geom_smooth(data=subset(weeklySoil[14:28, ]), method = "lm", formula = y ~ poly(x, 2), se = F) +
  # geom_smooth(aes(group = 1), method = "lm", formula = y ~ poly(x, 2)) +
  # stat_smooth(data=subset(weeklySoil[4:39, ]), method = "lm", formula = y ~ poly(x, 2), se = F) 
  # stat_smooth(method = "lm", formula = y ~ x, se=FALSE)
  geom_text_repel(data=subset(bulkDegDF, Transect == "Degraded (Bulk)"),
                  mapping = aes(y=B.mean.bulk, x= Date.ti, label=Week),
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.9, 
                  point.padding = unit(0.9, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05, show.legend = F)

Bsoil1

colnames(bulkDegDF)[which(names(bulkDegDF) == "Transect")] <- "Catchment"
limits_bulkdegCAT <- aes(ymin=B.min.bulk, ymax=B.max.bulk, x = Date.ti, colour = Catchment, group = Catchment)
solidpts <- ggplot() +
  # geom_point(size = 2) +
  # ggplot(data = wSoilBulkDeg, aes(Date.ti, B.mean.bulk, colour = Transect, shape=Transect, group = Transect)) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, B.mean.bulk, colour = Catchment, shape=Catchment, group = Catchment), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, PersistPrct, colour = RemainLabel, shape=RemainLabel, group = RemainLabel), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, DischPrct, colour = DischLabel, shape=DischLabel, group = DischLabel), size = 2) +
  geom_point(data = bulkDegDF, 
             aes(Date.ti, LeachPrctCorr, colour = LeachLabel, shape=LeachLabel, group = LeachLabel), size = 2) +
  
  geom_errorbar(data = bulkDegDF, limits_bulkdegCAT) + # With 2 data frames
  stat_smooth(data=subset(bulkDegDF, Catchment == "Degraded (Bulk)"), 
              mapping = aes(y = B.mean.bulk, x = Date.ti), # With 2 data frames
              colour = " black",
              method = "lm", formula = y ~ poly(x, 2), se=F) +
  
  stat_smooth(data=subset(bulkDegDF, RemainLabel == "Persistent frac. (Top Soil 1cm)"), 
              mapping = aes(y = PersistPrct, x = Date.ti), # With 2 data frames
              colour = "darkblue",
              method = "lm", formula = y ~ poly(x, 2), se=F) +
  #scale_colour_manual(values=c(" black" ,"#F8766D", "#7CAE00", "#619CFF")) +
  scale_colour_manual(values=c("black" , "chartreuse4",  "orangered2",  "dodgerblue4")) +
  scale_shape_manual(values = c(15, 18,  16, 17)) + 
  theme_bw() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        #axis.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        #axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  # xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  geom_text_repel(data=subset(bulkDegDF, Catchment == "Degraded (Bulk)"),
                  mapping = aes(y=B.mean.bulk, x= Date.ti, label=Week),
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.9, 
                  point.padding = unit(0.9, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05, show.legend = F)

solidpts

#colnames(wSoil)[which(names(wSoil) == "Transect")] <- "Transect-Deg."
emptypts <- 
  ggplot() +
  # geom_point(size = 2) +
  # ggplot(data = wSoilBulkDeg, aes(Date.ti, B.mean.bulk, colour = Transect, shape=Transect, group = Transect)) +
  geom_point(data = wSoil, 
             aes(Date.ti, B.mean.comp, colour = Transect , shape=Transect , group = Transect ), size = 2) +

  #scale_colour_manual(values=c(" black" ,"#F8766D", "#7CAE00", "#619CFF")) +
  
  scale_colour_manual(values=c("#F8766D","#7CAE00",  "#619CFF")) +
  scale_shape_manual(values = c(23,  6,  13)) +
  # guides(guide_legend(title = waiver()))+
  # labs(color = "Number of gears") +
  #scale_colour_manual(values=c("black" , "chartreuse4",  "orangered2",  "#F8766D", "dodgerblue4", "#619CFF", "#7CAE00")) +
  #scale_shape_manual(values = c(15, 18, 16,  23, 17, 13, 6)) + 
  theme_bw() +
  ylab("S-metolachlor Sinks (%)") +
  theme(#legend.position = "top",
        # legend.title = element_blank(),
        #axis.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank()
        #axis.text.x=element_text(angle = 45, hjust = 1)
        ) +
  # xlab("Date") +
  scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d"))

emptypts
#colnames(wSoil)[which(names(wSoil) == "Transect-Deg.")] <- "Transect"

## Merging both figures

# massBalNoL <- massBal + theme(legend.position='none')
BsoilNoL <- Bsoil1 + theme(legend.position='none')
MBalNoL <- massBalTop + theme(legend.position = 'none')

legend_deg_solid = get_legend(solidpts + 
                          theme(legend.title = element_text(face = "bold"), 
                                legend.text = element_text(size = 9), 
                                legend.key.height = unit(x = 0.35, units = 'cm'),
                                legend.background = element_rect(colour = "black")) + 
                          guides(colour = guide_legend(title.position = "top", ncol = 1, byrow = T)))

legend_deg_empty = get_legend(emptypts + 
                          theme(legend.title = element_text(face = "bold"),
                                legend.text = element_text(size = 9), 
                                legend.key.height = unit(x = 0.35, units = 'cm'),
                                legend.background = element_rect(colour = "black")) + 
                          guides(colour = guide_legend(title.position = "top", ncol = 1, byrow = T)))

legend_mass = get_legend(massBalTop + 
                           theme(legend.text = element_text(size = 9), 
                                 legend.key.height = unit(x = 0.4, units = 'cm'),
                                 legend.background = element_rect(colour = "black"),
                                  legend.title = element_blank()))
                         #+   guides(colour = guide_legend(title = "Mass Distribution",  title.position = "top", ncol =1, byrow = T)))

gridPartB <- plot_grid(BsoilNoL, MBalNoL,
                    ncol =1, nrow = 2, align ="v",
                    labels = c("A", "B"))
gridPartB

fig2 <- ggdraw() +
  draw_plot(gridPartB, x=0, y=0, width =  0.93, height = 1) + 
  draw_plot(legend_deg_solid, x=0.48, y =0.62, width = 0.70, height = 0.20) +
  draw_plot(legend_deg_empty, x=0.52, y =0.65, width = 0.75, height = 0.35) +
  draw_plot(legend_mass, x=0.5, y=0.05, width = 0.63, height = 0.28)

fig2

# ggsave(fig2, filename = "BvsMassBal.tiff", height = 17, width = 17.8, units = 'cm')
}

```

### Degradation per transect (no bulk)

```{r}

if (SHOW) {
  Bsoil2 = ggplot(wSoil, aes(x=Date.ti, y=B.mean.comp, colour=Transect, shape=Transect, group = Transect)) +
    # Bsoil2 = ggplot(wSoilBulkDeg, aes(x=Date.ti)) +
    # 
    # stat_smooth(aes(y=B.mean.bulk), method = "lm", formula = y ~ poly(x, 2), se=T) +
    # geom_point(aes(y=B.mean.com, colour=Transect, group = Transect)) +
    # geom_point(aes(y=B.mean.bulk, group=BulkLabel, colour="Bulk Isotopes")) +
    geom_point(size = 2) +
    
    stat_smooth( method = "lm", formula = y ~ poly(x, 2), se=F) +
    
    scale_colour_manual(values=c("#F8766D", "#7CAE00", "#619CFF")) +
    scale_shape_manual(values = c(15, 16, 17)) +
    # scale_shape_manual(values = c(23, 15, 16, 17)) +
    # geom_point(aes(x=Date.ti, y=B.comp, colour=Transect, group = Transect)) +
    # Theme and axes
    theme_bw() +
    ylab("Degr. %") +
    theme(legend.position = "top",
          #axis.title = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_blank()
          #axis.text.x=element_text(angle = 45, hjust = 1)
          ) +
    # xlab("Date") +
    scale_x_datetime(breaks = date_breaks("2 weeks"), labels = date_format("%b %d"))
    # ylab(expression(paste("Degradation %"))) +
    
    # scale_y_continuous(breaks = c(25, 50, 75, 100), limits = c(0, 100) ) +
    
    # geom_smooth(data=subset(weeklySoil[14:28, ]), method = "lm", formula = y ~ poly(x, 2), se = F) +
    # geom_smooth(aes(group = 1), method = "lm", formula = y ~ poly(x, 2)) +
    # stat_smooth(data=subset(weeklySoil[4:39, ]), method = "lm", formula = y ~ poly(x, 2), se = F) 
    # stat_smooth(method = "lm", formula = y ~ x, se=FALSE)
  
  
    #geom_text_repel(aes(y=B.mean.bulk, label=Wnum, color = factor(Transect)),
     #               size = 3,
      #              arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
       #             force = 0.5, 
        #            point.padding = unit(0.5, 'lines'), 
         #           max.iter = 2e3,
          #          nudge_x = .05)
  
  Bsoil2
}
```


### XY-Plots

```{r}
QC <- ggplot(AOdf, aes(y=AveDischarge.m3.h, x=Conc.mug.L, group = WeekSubWeek, color = Weeks)) +
  geom_point(size = 2) +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  theme(legend.title=element_blank()) +
  theme(plot.margin = unit(c(0,0.5,0,0), "lines")) +
  #stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  theme_bw() +
  theme(legend.position="none") +
  #scale_y_continuous(trans=log_trans(), breaks=c(1, 5, 10, 50, 100, 200)) +
  scale_x_continuous(trans=log_trans(), breaks=c(1, 3, 5, 10, 20)) +
  ylab(expression(paste("Ave. Discharge ", {(m^{3} / samp.hrs. )}))) +
  xlab(expression(paste("Conc. S-Meto.  ", {({mu}*g / L)}))) +
  geom_text_repel(aes(label=WeekSubWeek, color = factor(Weeks)),
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05)

QC

QD <- ggplot(AOdf, aes(y=AveDischarge.m3.h, x=diss.d13C, group = WeekSubWeek, color = Weeks)) +
  geom_point(size = 2) +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  theme(plot.margin = unit(c(0,0.8,0,0), "lines")) +
  #theme(legend.title=element_blank()) +
  #theme(legend.text = element_text(size = 10)) +
  theme(legend.position="none") +
  #stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  #scale_y_continuous(trans=log_trans(), breaks=c(1, 3, 5, 8, 10, 30, 50, 80, 100, 300)) +
  ylab(expression(paste("Ave. Discharge ", {(m^{3} / sample)}))) +
  ylab("") +
  scale_x_continuous(breaks=seq(-31.5, -26.5, 1)) +
  xlab(expression(paste({delta}^"13","C", ' (\u2030)'))) +
  geom_text_repel(aes(label=WeekSubWeek, color = factor(Weeks)),
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05)

QD

acd = plot_grid(QC, QD, ncol = 2, nrow = 1, align = "h")
acd

#ggsave(acd, filename = "Disch_Conc_Delta_XYlabs.png", width = 8, height = 5, units = "in", scale = 1)
#ggsave(acd, filename = "Disch_Conc_Delta_XY.png", width = 8, height = 5, units = "in", scale = 1)
#ggsave(acd, filename = "Disch_Conc_Delta_W.pdf", width = 8, height = 4.6, units = "in", scale = 1)
```





