---
title: "Data Analysis Prep"
author: "PAZ"
date: "8 ao?t 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
Sys.setlocale("LC_ALL", "English")
```

## Files

Imports: 

- **WeeklyHydroContam_R.csv** (water, Book 5)
- **MassBalance_R.csv** (soils, Book 7)

Generates (by merging above):

- **WaterSoils_R.csv**

## Import packages

```{r }

library("ggplot2")
library("ggrepel")
library("zoo")

# Melting data sets & changin axes
library("reshape2")

library("MASS")
```


# Lab parameters

```{r}
# Initial signature measured in tank
initialDelta = -32.253
```


## References

Modified from: D. Borcard & F. Gillet
Multivariate Analysis in Community Ecology: Constrained ordination and other analysis

Adapted from: Gwena?l Imfeld, LyGeS,2009


## Import and merge water and soil data sets


```{r}

# Check working directory
getwd()
# setwd("D:/Documents/these_pablo/Rscripts/Clustering")

waters = read.csv2("Data/WeeklyHydroContam_R.csv")



waters$ti <- as.POSIXct(strptime(waters$ti, "%Y-%m-%d %H:%M", tz="EST"))
colnames(waters)[colnames(waters) == "ti"] <- "Date.ti"
waters$Events <- factor(waters$Events, levels = unique(waters$Events))
waters$Event <- factor(waters$Event, levels = unique(waters$Event))

#colnames(waters)

# Integrate Catchment's Bulk signature for normalization of discharge signatures
# Soils
soils = read.csv2("Data/MassBalance_R.csv", 
                       na.strings=c('#DIV/0!', '', 'NA'), header = TRUE)
colnames(soils)[colnames(soils) == "ti"] <- "Date.ti"
soils$Date.ti <- as.POSIXct(strptime(soils$Date.ti, 
                                          "%Y-%m-%d %H:%M", tz="EST")) # csv typos, option 1
sum(is.na(soils$Date.ti)) == 0

#colnames(soils)

keepSoil <- c("WeekSubWeek", "Event", 
              "comp.d13C.SE.North", "comp.d13C.SE.Talweg", "comp.d13C.SE.South", 
              "f.max.comp", "f.mean.comp", "f.min.comp", "ngC.SD","ngC.SE", "N_compsoil", "N_ngC")
soils <- soils[ , !(names(soils) %in% keepSoil)]

names(soils)

repeatsInWater <- names(waters)[(names(waters) %in% names(soils))] # Not unique to waters
repeatsInWater <- repeatsInWater[2:length(repeatsInWater)]
waters <- waters[ , !(names(waters) %in% repeatsInWater)]

watSoilMerged <- merge(waters, soils, by = "Date.ti", all = T)

if (is.na(watSoilMerged$WeekSubWeek[nrow(watSoilMerged)])) {
  watSoilMerged <- watSoilMerged[1:nrow(watSoilMerged)-1, ]
}
# d13obs <- waters$SD.d13C
# totd13obs <- length(d13obs) - sum(is.na(d13obs))

watSoilMerged$DD13.Bulk <- watSoilMerged$BulkCatch.d13-initialDelta 

colnames(watSoilMerged)


```

## Reduce variable size 

```{r}

dropNoUse <- c("Markers" , "TimeDiff", 
               "se.d13C", "MES.mg.L", "MES.sd", "MO.mg.L", "filt.se.d13C", "f.diss", "f.filt",
               "Appl.Mass.g", 
               "FracDiss", "FracFilt",
               # Nanogram obs numbers.
               "N_ngC.diss", "N_ngC.fl", 
               # Remove in mg units 
               "DissSmeto.mg", "DissSmeto.mg.SD" ,
               "DissOXA.mg", "DissOXA.mg.SD",
              "DissESA.mg", "DissESA.mg.SD" , 
              "FiltSmeto.mg", "FiltSmeto.mg.SD" ,
              "TotSMout.mg","TotSMout.mg.SD",
              # Transect areas
              "ID.N", "ID.T", "ID.S", "Area.N", "Area.T", "Area.S",
              # Standard errors
              "se.d13C", "filt.se.d13C",
              # Degradation 
              "B.diss.x", "B.filt.x",
              "B.mean.comp.North", "B.max.comp.North" , "B.min.comp.North",
              "B.mean.comp.Talweg", "B.max.comp.Talweg", "B.min.comp.Talweg",
              "B.mean.comp.South", "B.max.comp.South" , "B.min.comp.South", 
              "FracDeltaOut" ,
              # Isotopes (DD already included)
               "comp.d13C.North","comp.d13C.SD.North", 
              "comp.d13C.Talweg", "comp.d13C.SD.Talweg" #,
              # Masses on transects
              # "MassSoil.g.North", "MassSoil.g.Talweg", "MassSoil.g.South" 
              )

watSoilMerged <- watSoilMerged[ , !(names(watSoilMerged) %in% dropNoUse)]


# Date conversion correct: 
sum(is.na(watSoilMerged$Date.ti)) == 0
colnames(watSoilMerged)
sum(is.na(watSoilMerged$maxQ))
```


## Variable generation  

We would like to determine whether there are different clusters in the data.

Via response variables:

- Concentrations ($\mu g / L$)
- MEL-sm (g)
- Loads (SM g)
- Transformation products (OXA and ESA in $\mu g / L$ and in loads $g$)

Via hydrological characteristics:

- Event index: 
$$ \frac{I_{max} \cdot R_{tot} } {D} $$

- Event duration ($t_f - t_i$)
- Volume discharged ($\sum^N_{i=1}Q_i \cdot dt_i$ , N: no. of measurements within the event)
- Average discharge ($\sum^N_{i=1}Q_i/N$)

Imax = max rainfall intensity mm/h ; 
Rtot = rainfall amount (mm); 
D = duration (min)

"A high EVI represents a short but intense rainfall event, whereas a low EVI indicates an event with a low intensity but long duration. The catchment response time is defined as the time between the gravity centre of the rain event and the peak outflow. (Baartman et al., 2013; in Lefrancq etal2017)"

The EVI has been adapted to reflect discharge index such that:

- Discharge index A [m3/h x m3/h]
$$DIa = \frac{Q_{max} \cdot V_{tot} }{D}  $$  


```{r}

watSoilMerged$DIa <- watSoilMerged$maxQ*watSoilMerged$Volume.m3/watSoilMerged$Duration.Hrs
# watSoilMerged$DIb <- watSoilMerged$Volume.m3/watSoilMerged$Duration.Hrs * 1/watSoilMerged$maxQ

```

## Normalization by remaining S-met mass and bulk (\\$\Delta \delta$)

This section normalizes outlet response variables (S-met and TPs loadings and $\Delta \delta$ values).

1. Convert date-time to cumuative days since first measurements (needed for step 2)
2. Compute linear models to predict bulk catchment soils $\Delta \delta$ and mass remaining. 
3. Normalize outlet $\Delta \delta$ and S-met loadings by predicted values.  

```{r}

watSoilMerged$CumSmpDays <- cumsum(watSoilMerged$Sampled.Hrs)/24
names(watSoilMerged)

DD.glm1 <- glm(watSoilMerged$DD13.Bulk ~ watSoilMerged$CumSmpDays)
# DD.glm2 <- glm(watSoilMerged$DD13.Bulk ~ (watSoilMerged$CumSmpDays)^2, family = gaussian)
summary(DD.glm1) # Both models were identical

DDdiss.glm1 <- glm(watSoilMerged$DD13C.diss ~ watSoilMerged$CumSmpDays)

# Input isotope predicted values 
watSoilMerged$DD.Bulk.imp <- predict(DD.glm1, data.frame(watSoilMerged$CumSmpDays))
watSoilMerged$DD.diss.imp <- predict(DDdiss.glm1, data.frame(watSoilMerged$CumSmpDays))
View(watSoilMerged[, c(1, 119, ncol(watSoilMerged))])
# Input linear approx mass soils
watSoilMerged$CatchMassSoil.g.imp <- na.approx(watSoilMerged$CatchMassSoil.g)


# Isotopes Bulk Soils
ggplot(data = watSoilMerged, aes(x = CumSmpDays))+
  geom_point(aes(y = DD13.Bulk)) +
  geom_point(aes(y = DD.Bulk.imp)) +
  stat_smooth(aes(x = CumSmpDays, y = DD13.Bulk ), method = "lm", formula = y ~ poly(x, 2), se = T, alpha = 0.2, size=0.2) +
  theme_bw()


ggplot(data = watSoilMerged, aes(x = CumSmpDays))+
  geom_point(aes(y = CatchMassSoil.g.imp), col = "forestgreen") +
  geom_point(aes(y = CatchMassSoil.g), col = "red") +
  theme_bw()

# View(watSoilMerged[, c(1:29, 47:60, 70, 102: ncol(watSoilMerged))])

# Option 1. 
# Divide by estimated mass in catchment available # [-]
# watSoilMerged$SM.g.nrm <- watSoilMerged$TotSMout.g/watSoilMerged$BalMassDisch.g # [-]
# watSoilMerged$MEL.g.nrm <- watSoilMerged$MELsm.g/watSoilMerged$BalMassDisch.g # [-]

# Option 2
# Divide by estimated prct. mass in catchment available # [g]
# watSoilMerged$CumPrctMassOut <- cumsum(watSoilMerged$prctMassOut)
# watSoilMerged$SM.g.nrm.prc <- watSoilMerged$TotSMout.g/watSoilMerged$CumPrctMassOut # [-]
# watSoilMerged$MEL.g.nrm.prc <- watSoilMerged$MELsm.g/watSoilMerged$CumPrctMassOut # [-]


watSoilMerged$SM.g.nrm <- watSoilMerged$TotSMout.g/watSoilMerged$CatchMassSoil.g.imp # [-]
watSoilMerged$TP.g.nrm <- (watSoilMerged$MELsm.g-watSoilMerged$TotSMout.g)/watSoilMerged$CatchMassSoil.g.imp # [-]
watSoilMerged$DD.diss.nrm <-  watSoilMerged$DD13C.diss/watSoilMerged$DD.Bulk.imp # Losses values where no BULK was obtained
watSoilMerged$DD.diss.nrmSD <- watSoilMerged$SD.d13C/watSoilMerged$DD.Bulk.imp
# watSoilMerged$DD.diff.prc <- ifelse( is.na(watSoilMerged$DD13C.diss) , NA,
#                                    ((watSoilMerged$DD13C.diss-watSoilMerged$DD.Bulk.imp)/watSoilMerged$DD.Bulk.imp)*100)

names(watSoilMerged)

#View(watSoilMerged[, c( 1, 32 , 73, 81:83 )])

```

## Variable reduction for analysis 

```{r}

waterXY <- watSoilMerged

write.csv2(waterXY, 
           'Data/WaterSoils_R.csv', row.names = F)
sum(is.na(watSoilMerged$maxQ))

# waterXY <-  waterSmall[ , (names(waterSmall) %in% hydro)]
# waterXY.nona <- waterXY[complete.cases(waterXY), ]

```


## Rainfall by period

## Dry periods revised based on rainfall (not discharge)

```{r}

# Merge different cummulative rains (ignore for now)
if (FALSE){
  rain30 <- read.csv2("Data/30minRain.csv")
  rain7 <- read.csv2("Data/7hrRain.csv")
  rain12 <- read.csv2("Data/12hrRain.csv")
  
  formTime <- function(x, newColName){
    x$Time <- as.POSIXct(strptime(x$Time, "%d/%m/%Y %H:%M", tz="EST"))
    colnames(x)[colnames(x) == "Cumm.mm"] <- newColName
    colnames(x)[colnames(x) == "Time"] <- "Date.ti"
    return(x)
  }
  
  rain30 <- formTime(rain30, "Cum30min")
  rain7 <- formTime(rain7, "Cum7hr")
  rain12 <- formTime(rain12, "Cum12hr")
  
  rain <- Reduce(function(...) merge(..., all=TRUE), list(rain30, rain7, rain12))
  
  watRain <- merge(watSoilMerged, rain12, by = "Date.ti", all = T)
  watRain <- Reduce(function(...) merge(..., all=TRUE), list(watSoilMerged, rain7, rain12))
  
  } else {
    watRain <- watSoilMerged
}


```

## Reducing variables

```{r}
names(watSoilMerged)
```



