---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

## Purpose

This file prepares data sets for predictive analysis. Several machine learning tools are evaluated against their ability to predict normalized  dependent variables ($y_i$) based on event characterisitics ($x_i$). Normalization is done by dividing the bulk isotope signatures or bulk mass measured in soil transects at of the nearest soil sampling period.

Dependent (normalized) variables include:

- Signatures 
- SM masses 
- TPs masses
- MEL$_{SM}$ (Mass equivalent loads for S-metolachlor)


Independent variables include:

- Event duration ($t_f - t_i$)
- Volume discharged ($\sum^N_{i=1}Q_i \cdot dt_i$, N: no. of measurements within the event)
- Average discharge ($\sum^N_{i=1}Q_i/N$)

- Kurtosis (Tailed-ness)
- Skweness (symmetry extent)

Imports: 

- **.csv** 

Generates:

- ****

## Required R-packages:

```{r, message=FALSE}

library("plyr")
library("dplyr")

```

## Working directory

```{r, message=FALSE}

getwd()

```

## Models to evalaute

# CART - Regression Trees (RTs)

- In regression trees the output of each leaf is a real number (dependent variable), namely the average of the value at that leaf
- Unlike linear regression models, RTs can capture non-linearities
- Compare fitness based on "cp" parameter
- Apply cross-validation

