---
title: "CSIA Monitoring Scope" 
author: "PAZ" 
date: "22/09/2017"
output: pdf_document
---


```{r, echo=FALSE, message=FALSE, include=FALSE} 
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = F)

Sys.setlocale("LC_ALL", "English") 

library("tidyr")
#library("dplyr")
library("cowplot")
library("gridExtra")
library("ggplot2")
library("ggrepel")
library("zoo")
library("scales")
#library("plotly")
library("Cairo")
```

## Purpose

Plot sampled cocentrations agains theoretical minimum mass vs. volume relations for multi-element CSIA. 

Data file:

- **MonitoringScope_R.csv** (R generated, Book 5 $\approx$ Line 350 )

```{r} 
outscope = read.csv2('Data/MonitoringScope_R.csv', header = T) 
#names(outscope)

out = outscope[, c("WeekSubWeek" , 
                   #"AveDischarge.m3.h", 
                   "Sampled", "Vol.SPE.L", "Conc.mug.L", "DD13C.diss")]
out$CSIA = ifelse(is.na(out$DD13C.diss), "No", "Yes")
out$DD13C.diss = NULL
out$Conc.mug.L = round(out$Conc.mug.L, digits = 2)
out$C.txt = sprintf("%.3f", round(out$Conc.mug.L, digits = 3) ) 
out$Conc.mug.L  = NULL
colnames(out)[colnames(out) == "Vol.SPE.L"] <- "Field.Vsmp"
```



```{r}
# Theoretical Concentrations
iterations = 6000
variables = 1
conc = 0.00
C.smp = matrix(ncol = variables, nrow = iterations)

for (i in 1:iterations){
  C.smp[i,] = conc
  if (conc < 0.05){
    conc = conc + 0.001
  } else {
    conc = conc + 0.010  
  }
}
```



# Carbon

## Conversion to ng

For a given sample of concentration $C$ [$\mu g/L$], considering a 4
$\mu L$ injection and a vial volume $x_{vial}$ of 500 $\mu L$:

$$ C~[\mu g/L] \cdot \frac{10^3~ng}{1~\mu g}
\cdot \frac{1~mol_{S-met}}{283.796~g} 
\cdot \frac{12.0107~g}{1~mol_{C}} 
\cdot 15 
\cdot \frac{4~[\mu L~inj.]}{x_{vial}~[\mu L~vial]} 
\cdot X_{smp}~[L] = M_{C}$$

The same equation re-arranged to compute it numerically so as to yield 10 $ng$ of carbon, from a 500 $\mu L$ vial:

$$ C~[\mu g/L] \cdot \frac{10^3~ng}{1~\mu g}
\cdot \frac{1~mol_{S-met}}{283.796~g} \cdot \frac{12.0107~g}{1~mol_{C}} \cdot 15 \cdot \frac{4~[\mu L~inj.]}{500~[\mu L~vial]} \cdot \frac{1}{10~ ng_{C}} = \frac{1}{X_{smp}~[L]}$$

```{r, echo=FALSE }
df.C = data.frame(C.smp)

# Early assumption, min. Carbon 15 ng:
# minMass = 15
# Late correction:
minMass = 10
elementMass = 12.0107 # g/mol
atoms = 15

# 100 used for GC, TPs and each isotope, plus 1x backup = 500 uL
# For re-concentration steps:
# 200uL (out of 500) are considered for GC and TPs, there is no backup and 300uL is reconcentrated to either:
# 100uL (yielding two samples) or 30uL (yielding one sample)
df.C$uL500C.Vsmp = 1/(df.C$C.smp*(10^3)*(elementMass/283.796)*atoms*(4/500)*(1/minMass))
df.C$uL50C.Vsmp = 1/(df.C$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/100)*(1/minMass)) 
df.C$uL30C.Vsmp = 1/(df.C$C.smp*(10^3)*(elementMass/283.796)*atoms*(350/500)*(4/30)*(1/minMass)) # 100 used for GC

df.C = df.C[-1,]
df.C$C.txt = sprintf("%.3f", round(df.C$C.smp, digits = 3) )
```


# Chlorine

$$ C~[\mu g/L] \cdot \frac{10^3~ng}{1~\mu g}
\cdot \frac{1~mol_{S-met}}{283.796~g} \cdot \frac{35.453~g}{1~mol_{Cl}} \cdot 1 \cdot \frac{4~[\mu L~inj.]}{500~[\mu L~vial]} \cdot \frac{1}{10~ ng_{Cl}} = \frac{1}{X_{smp}~[L]}$$

```{r}
df.Cl = data.frame(C.smp)

minMass = 10
elementMass = 35.453
atoms = 1
# 0.008, 0.06, 0.1
# 100 used for GC, TPs and each isotope, plus 1x backup = 500 uL
df.Cl$uL500Cl.Vsmp = 1/(df.Cl$C.smp*(10^3)*(elementMass/283.796)*atoms*(4/500)*(1/minMass))
df.Cl$uL50Cl.Vsmp = 1/(df.Cl$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/100)*(1/minMass)) 
df.Cl$uL30Cl.Vsmp = 1/(df.Cl$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/60)*(1/minMass)) # 100 used for GC

df.Cl = df.Cl[-1,]

df.Cl$C.txt = sprintf("%.3f", round(df.Cl$C.smp, digits = 3) )
```

# Nitrogen

$$ C~[\mu g/L] \cdot \frac{10^3~ng}{1~\mu g}
\cdot \frac{1~mol_{S-met}}{283.796~g} \cdot \frac{14.007~g}{1~mol_{N}} \cdot 1 \cdot \frac{4~[\mu L~inj.]}{500~[\mu L~vial]} \cdot \frac{1}{30~ ng_{N}} = \frac{1}{X_{smp}~[L]}$$

```{r}
df.N = data.frame(C.smp)

minMass = 30 # ng
elementMass = 14.0067 # g/mol
atoms = 1

# 100 used for GC, TPs and each isotope, plus 1x backup = 500 uL
df.N$uL500N.Vsmp = 1/(df.N$C.smp*(10^3)*(elementMass/283.796)*atoms*(4/500)*(1/minMass))
df.N$uL50N.Vsmp = 1/(df.N$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/100)*(1/minMass)) 
df.N$uL30N.Vsmp = 1/(df.N$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/60)*(1/minMass)) # 100 used for GC

df.N = df.N[-1,]

df.N$C.txt = sprintf("%.3f", round(df.N$C.smp, digits = 3) )
```

# Hydrogen

$$ C~[\mu g/L] \cdot \frac{10^3~ng}{1~\mu g}
\cdot \frac{1~mol_{S-met}}{283.796~g} \cdot \frac{1.00794~g}{1~mol_{H}} \cdot 15 \cdot \frac{4~[\mu L~inj.]}{500~[\mu L~vial]} \cdot \frac{1}{30~ ng_{N}} = \frac{1}{X_{smp}~[L]}$$

```{r}
df.H = data.frame(C.smp)

minMass = 30 # ng
elementMass = 1.00794 # g/mol
atoms = 22

# 100 used for GC, TPs and each isotope, plus 1x backup = 500 uL
df.H$uL500H.Vsmp = 1/(df.H$C.smp*(10^3)*(elementMass/283.796)*atoms*(4/500)*(1/minMass))
df.H$uL50H.Vsmp = 1/(df.H$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/100)*(1/minMass)) 
df.H$uL30H.Vsmp = 1/(df.H$C.smp*(10^3)*(elementMass/283.796)*atoms*(300/500)*(4/60)*(1/minMass)) # 100 used for GC

df.H = df.H[-1,]

df.H$C.txt = sprintf("%.3f", round(df.H$C.smp, digits = 3) )
```


```{r}
## Merging dataframes
df.Cl$C.smp = NULL
df.N$C.smp = NULL
df.H$C.smp = NULL
dfout.CCl = merge(df.C, df.Cl, by = "C.txt", all = T)
dfout.CClN = merge(dfout.CCl, df.N, by = "C.txt", all = T)
dfout.CClNH = merge(dfout.CClN, df.H, by = "C.txt", all = T)
dfout.CClNH = merge(dfout.CClNH, out, by = "C.txt", all = T)



# dfout.CCl$AveDischarge.m3.h = NULL

# Melting dataframe
df.tidy <- dfout.CClNH %>%
  gather(measure, value, 
         -C.smp, 
         -C.txt,
         -WeekSubWeek,
         #-AveDischarge.m3.h, 
         -Sampled, -CSIA) %>% # Melts data frame
  separate(measure, into = c("VialSize", "temporary_var")) %>% # parses the sep = "." into...
  # VialSize will be first string of variable name
  spread(temporary_var, value)
  
```

## Plotting

### Multi element CSIA
```{r, width = 10, height = 6}
#View(df.tidy.CClN)
size = 4
g = ggplot(data= df.tidy)+
  geom_line(data=subset(df.tidy, VialSize != "Field" & 
                          (VialSize == "uL500C" | 
                             VialSize == "uL500N" |
                               VialSize == "uL500Cl" |
                               VialSize == "uL500H" |
                             VialSize == "uL30C")), 
            aes(x = Vsmp, y = C.smp, group = VialSize, colour = VialSize) ) +
  geom_point(data=subset(df.tidy, VialSize == "Field" & !is.na(CSIA)), 
             colour = "blue", alpha = 0.5, size = size,
             aes(x = Vsmp, y = C.smp, 
                 group = VialSize,  shape = CSIA #, 
                 #size = AveDischarge.m3.h
                 )) +
  #geom_line(data=subset(df.tidy, VialSize == "uL30")) + 
  #geom_line(data=subset(df.tidy, VialSize == "uL50")) +
  #geom_line(data=subset(df.tidy, VialSize == "uL500")) +
  theme_minimal() +
  theme(#legend.position = "bottom",
        text = element_text(size=17)) +
  scale_y_continuous(trans=log_trans() , breaks=c(0.01, 0.05, 1, 5, 15, 30, 60) )+
  # scale_x_continuous( limits = c(0,40) , expand = c(0,0) )
  coord_cartesian(xlim=c(0,15)) + 
  scale_color_manual(name= "MB + Feasible CSIA", 
                     # Actual order:
                     # "uL30C", "uL500C", "uL500Cl", "uL500H", "uL500N"
                      values = c("#F8766D", "#00BA38", "#C77CFF", "#619CFF", "#DE8C00"
                                 # Blue: "#619CFF"
                                 # Brown: "#DE8C00"
                                 #c("black", "#F8766D", "#00BFC4", "#DE8C00", 
                                 #"black", "#D55E00",  "#00BFC4",  "#B79F00", "#00BA38"
                                 # Bulk, North, outlet, South, Valley
                                 #"#D55E00", "darkgreen", "dodgerblue"
                                 ),
                     breaks=c("uL500N" , 
                              "uL500H",
                              "uL500Cl",
                              "uL500C" , 
                              "uL30C"),
                     labels=c("N + H + Cl + C" , 
                              "H + Cl + C" , 
                              "Cl + C", 
                              "C",  
                              "Cx")
                     ) + 
  xlab(expression(paste( "Volume ", "(L)"))) +
  ylab(expression(paste("Outlet S-met Conc.  ", {({mu}*g / L)}))) +
  annotate("rect", xmin=3.9, xmax=4, ymin=0, ymax=30, alpha=1) +
  geom_segment(aes(x = 5, y = 0.005, xend = 4.1, yend = 0.005),
                   arrow = arrow(length = unit(1/2, 'picas'), type = "closed")) +
  annotate("text", 
           x = 8, y = 0.005, 
           label= "Weekly sampler capacity", parse=F, size = size) +
  

  annotate("text", 
           x = 5, y = 80, 
           label= "> Triple CSIA \nn = 0%", parse=F, size = size) +
  geom_segment(aes(x = 5, y = 40, xend = 5, yend = 13),
                   arrow = arrow(length = unit(1/2, 'picas'), type = "closed")) + 
  annotate("text", 
           x = 7.5, y = 30, 
           label= "Double CSIA \nn = 18%", parse=F, size = size) +
  geom_segment(aes(x = 7.5, y = 15, xend = 7.5, yend = 2),
                   arrow = arrow(length = unit(1/2, 'picas'), type = "closed")) +
  annotate("text", 
           x = 10.7, y = 20, 
           label= "C-CSIA \nn = 47%", parse=F, size = size) +
  geom_segment(aes(x = 10.7, y = 10, xend = 10.7, yend = 0.2),
                   arrow = arrow(length = unit(1/2, 'picas'), type = "closed")) +
  
  annotate("text", 
           x = 14, y = 25.0, 
           label= "C-CSIA \nn = 76% \n(extra re-conc. \nstep)", 
           parse=F, size = size) +
  geom_segment(aes(x = 14, y = 8, xend = 14, yend = 0.015),
                   arrow = arrow(length = unit(1/2, 'picas'), type = "closed")) 
  
# g

SAVE = F
PC = T
if (SAVE){
  if (PC){
    ggsave(g, 
       filename = "D:/Documents/these_pablo/WriteUp/Altek_PNAS_LaTeX_SI/images/Scope.pdf", 
       device = "pdf", dpi = 300,  scale = 1, # )# ,
       width = 8.7, height = 6)

  }
}
# ggsave(g, filename = "images/CSIAScope.png"  , width = 12, height = 8 )#, units = "cm", scale = 1)
# dfout.C$check500 = ifelse(dfout.C$uL500.Vsmp >= dfout.C$Field.Vsmp, 1, 0)
```

```{r}
library(scales) 
#show_col(hue_pal()(12))
```

## Plotting with labels

```{r}
ggplot(data= df.tidy.CClN)+
  geom_line(data=subset(df.tidy.CClN, VialSize != "Field" & 
                          ( # VialSize == "uL500Cl" | 
                             VialSize == "uL500C" | 
                             VialSize == "uL500N" |
                             VialSize == "uL30C" | VialSize == "uL30Cl")), 
            aes(x = Vsmp, y = C.smp, group = VialSize, colour = VialSize) ) +
  geom_point(data=subset(df.tidy.CClN, VialSize == "Field" & !is.na(CSIA)), 
             colour = "blue", alpha = 0.5,
             aes(x = Vsmp, y = C.smp, 
                 group = VialSize,  shape = CSIA #, 
                 #size = AveDischarge.m3.h
                 )) +
  #geom_line(data=subset(df.tidy, VialSize == "uL30")) + 
  #geom_line(data=subset(df.tidy, VialSize == "uL50")) +
  #geom_line(data=subset(df.tidy, VialSize == "uL500")) +
  theme_minimal() +
  scale_y_continuous(trans=log_trans(), breaks=c(0.01, 0.05, 1, 5, 10, 20, 30, 60) )+
  # scale_x_continuous( limits = c(0,40) , expand = c(0,0) )
  coord_cartesian(xlim=c(0,4)#, 
                  #ylim = c(0,60)
                  ) + 
  scale_color_manual(name= "MB + Feasible CSIA", 
                     # Actual order:
                     # 
                      values = c("#B79F00", "#F8766D", "#00BFC4", "#C77CFF", "#00BA38"
                                 # Blue: "#619CFF"
                                 # Brown: "#DE8C00"
                                 #c("black", "#F8766D", "#00BFC4", "#DE8C00", "#00BA38"
                                 #"black", "#D55E00",  "#00BFC4",  "#B79F00", "#00BA38"
                                 # Bulk, North, outlet, South, Valley
                                 #"#D55E00", "darkgreen", "dodgerblue"
                                 ),
                     breaks=c("uL500N" , "uL500C" , 
                              # "uL500Cl",  
                              "uL30C",  "uL30Cl"),
                     labels=c("N + Cl + C" , "Cl + C" , 
                              # "Cl", 
                              "C",  "Cl")
                     ) +
  geom_text_repel(data = subset(df.tidy.CClN, (!is.na(WeekSubWeek) & 
                                                 CSIA == "Yes" &
                                                 VialSize == "Field" &
                                                 C.smp >= 1 & 
                                                 Vsmp > 0.5
                                                 ) ), aes(x = Vsmp, y = C.smp, label=WeekSubWeek),
                 arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                 force = 1, 
                 point.padding = unit(1.0, 'lines'), 
                 max.iter = 2e3,
                 nudge_x = .2)
```

## Percentages

```{r}
# Total Samples
dfout.CClNH$NumSamp = ifelse(dfout.CClNH$Sampled == "Sampled", 1, 0)
numSamp = sum(dfout.CClNH$NumSamp, na.rm = T)

print("% Samples capable of triple CSIA")
dfout.CClNH$triple = ifelse(dfout.CClNH$uL500H.Vsmp <= dfout.CClNH$Field.Vsmp, 1, 0)
numTriple = sum(dfout.CClNH$triple, na.rm = T)
(numTriple/numSamp)*100 # 0 %

# Total capable of double CSIA
print('% Samples capable of double CSIA')
dfout.CClNH$double = ifelse(dfout.CClNH$uL500Cl.Vsmp <= dfout.CClNH$Field.Vsmp, 1, 0)
numDouble = sum(dfout.CClNH$double, na.rm = T)
(numDouble/numSamp)*100 # 18 %

print('% Samples capable of single C - CSIA')
dfout.CClNH$singleC = ifelse(dfout.CClNH$uL500C.Vsmp <= dfout.CClNH$Field.Vsmp, 1, 0)
numSingleC = sum(dfout.CClNH$singleC, na.rm = T)
(numSingleC/numSamp)*100 # 47 %

print('% Samples capable of single Cl - CSIA')
dfout.CClNH$singleCl = ifelse(dfout.CClNH$uL30C.Vsmp <= dfout.CClNH$Field.Vsmp, 1, 0)
numSingleCl = sum(dfout.CClNH$singleCl, na.rm = T)
(numSingleCl/numSamp)*100 # 76 %
```

