---
title: "SI Graphs Water Research"
author: "PAZ"
date: "23 novembre 2017"
output: pdf_document
---


```{r, echo=FALSE, message=FALSE, include=FALSE}
Sys.setlocale("LC_ALL", "English")
```

## Purpose

This produces a Hyetograph (rainfall and discharge data) based on 2 min intrevals. 

Input files:

- **groupAlteck2016_R** (Contains discharge and rainfall @ 2 min)

Output files:

- **nothing for now** 

## Libraries
```{r}
library(dplyr)
library(tidyr)
library(zoo)
library(reshape)
library(ggplot2)
library("ggrepel")

library("plotly")
library("cowplot")
library("gridExtra")
library("Cairo")
library("GGally")
library("scales")

library("plotKML")

# getwd()
# setwd("D:/Documents/these_pablo/Alteckendorf2016/HydrologicalMonitoring")
```

## Data

```{r}
# Discharge and 2 min rainfall 
hydro = read.csv2("Data/groupAlteck2016_R.csv")
hydro$Date = as.POSIXct(strptime(hydro$DateCheck.S,
                                 "%d/%m/%Y %H:%M", tz="EST") )
# names(hydro)
qra <- hydro[, c("Date", "Q.HW1", "Rain.mm")]

# Outlet Concentrations 
aodf = read.csv2("Data/WeeklyHydroContam_R.csv")
# names(aodf)
c = aodf[, c("ti", "Conc.mug.L" ,  "Conc.SD" , "Event", "Events")]
c$ti <- as.POSIXct(strptime(c$ti, "%Y-%m-%d %H:%M", tz="EST"))
sum(is.na(c$ti)) == 0

# Discharge and concentrations
ch = merge(qra, c, by.x = "Date", by.y="ti", all = T)

# Transect soils
weeklySoil = read.csv2("Data/WeeklySoils_Rng.csv", na.strings=c('#DIV/0!', '', 'NA'), header = TRUE)
weeklySoil$Date.ti <- as.POSIXct(strptime(weeklySoil$Date.ti, "%Y-%m-%d %H:%M", tz="EST")) 
sum(is.na(weeklySoil$Date.ti))
# weeklySoil$Conc.ComSoil.SD <- 
#  ifelse(weeklySoil$Conc.ComSoil.SD == as.character("#DIV/0!"), NA, as.numeric(as.character(weeklySoil$Conc.ComSoil.SD)))
                                  
# weeklySoil = weeklySoil %>%
#  group_by(Transect) %>%
#  arrange(Transect, Wnum)

# names(weeklySoil)
```

## Conversions

Convert rainfall data [$mm$] to the same units as discharge [$m^3/h$]

```{r}
ch$Rain.mm = ifelse(is.na(ch$Rain.mm), 0, ch$Rain.mm)
ch$Q.HW1 = ifelse(is.na(ch$Q.HW1), 0, ch$Q.HW1)

# Catchment area
area <- 47*10**4 # [m2]
ch$precip_m3 = ch$Rain.mm/10^3 * area

sum(is.na(ch$precip_m3))
sum(is.na(ch$Q.HW1))
# Calculate the range needed to avoid having your hyetograph and hydrograph overlap 
maxRange <- 1*(max(ch$precip_m3) + max(ch$Q.HW1)) - 800

# Create a function to backtransform the axis labels for precipitation
precip_labels <- function(x) {round( ((x / area) * 10^3), digits = 2) } # X will be precip_m3 -> convert to mm

lims <- as.POSIXct(strptime(c("2016-03-25 00:04:00","2016-07-12 01:00:00"), format = "%Y-%m-%d %H:%M"))    


# Plot the data
hy = ggplot(data = ch,
       aes(x = Date)) +
  theme_minimal() +
  # Use geom_tile to create the inverted hyetograph. geom_tile has a bug that displays a warning message for height and width, you can ignore it.
  geom_tile(aes(y = -1*(precip_m3/2-maxRange), # y = the center point of each bar, as geom_tile uses variable as center point
                height = precip_m3,
                width = 1
                ),
            fill = "gray50",
            color = "forestgreen") +
  theme(# axis.text.x=element_text(angle = 45, hjust = 1) 
        # axis.text.x=element_blank(),
        axis.title.x=element_blank()
        #legend.position="top"
        )+
  scale_x_datetime(limits = lims, breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  # Plot your discharge data
  geom_line(aes(y = Q.HW1),
            color = "blue") +
  # Create a second axis with sec_axis() and format the labels to display the original precipitation units.
  # ylab(expression(atop("Conc. S-met ", paste({({mu}*g / L)} )))) +
  scale_y_continuous(name = expression(atop("Discharge", paste({({m^3}/h)} ))),
                     sec.axis = sec_axis(trans = ~-1*(.-maxRange), # Equivalent to: y2 = -1*(y1 - maxRange) 
                                         name = expression(atop("Precipitation", paste("(mm)"))),
                                         labels = precip_labels)) # x here is = precip_cuft

hy
```



# Concentrations

```{r}
names(ch)
ch$Event = na.locf(ch$Event)
ch$Event = as.factor(ch$Event)
ch$Season = ifelse(ch$Date > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Late", "Early")

tags = FALSE

out_dates = ggplot(data = ch,
       aes(x = Date, y = Conc.mug.L)) +
  geom_errorbar(aes(ymin = Conc.mug.L - Conc.SD, ymax = Conc.mug.L + Conc.SD, colour = Season)) +
  geom_point(aes(group = Event, colour = Season)) +
  theme_minimal() +
  ylab(expression(atop("Conc. S-met ", paste({({mu}*g / L)} )))) +
  scale_y_continuous( breaks = c(28, 24, 20, 16, 12, 8, 4, 0), limits = c(0, 28) ) +
  
  # Smooth linear models
  geom_smooth(data=subset(ch[2203:43539, ]), method = "lm", formula = y ~ x, se = F,  linetype="dashed") +
  geom_smooth(data=subset(ch[43539:nrow(ch)-1000, ]), method = "lm", formula = y ~ x, se= F,  linetype="dashed") +
  
  scale_x_datetime(limits = lims, breaks = date_breaks("2 weeks"), labels = date_format("%b %d")) +
  guides(col = guide_legend(nrow = 2, title.position = "top"))
  
out = out_dates +   
  theme(# axis.text.x=element_text(angle = 45, hjust = 1) ,
        # axis.text.x=element_blank(),
        axis.title.x=element_blank(),
        legend.position="right"
        )
  

if (tags){
  out + geom_text_repel(aes(label=Events), # WeekSubWeek or Weeks
                  size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                  max.iter = 2e3,
                  nudge_x = .05, show.legend = F)
  
}
  
out_dates
```


## Soil Concentrations

```{r}
weeklySoil$Transect = as.character(weeklySoil$Transect)
weeklySoil$Transect = ifelse(weeklySoil$Transect == "N", "North", 
                             ifelse(weeklySoil$Transect == "T", "Valley",
                                    ifelse(weeklySoil$Transect == "S", "South", NA)
                             )
)
weeklySoil$Transect <- factor(weeklySoil$Transect, levels = c("North", "Valley", "South"))
levels( weeklySoil$Transect)

#, fig.height=3, fig.width=3}
lb1a2 <- paste("App.")
lbW012 <- paste("App.W0/1/2")
lbW9 <- paste("App.W9")

limits_conc_soil <- aes(ymin=Conc.mug.g.dry.soil-Conc.ComSoil.SD,  ymax=Conc.mug.g.dry.soil+Conc.ComSoil.SD)
#limits_conc_soil <- aes(ymin=mean-0.5, ymax=mean+0.5)

pd <- position_dodge(0.5) # move them .05 to the left and right

# xlims = xlim(as.POSIXct('2016-03-25 00:04:00', tz = 'EST'), as.POSIXct('2016-07-12 01:00:00', tz = 'EST') )

tags = FALSE

co = ggplot(weeklySoil[4:48, ], 
           aes(x=Date.ti, y=Conc.mug.g.dry.soil, colour=Transect, group = Transect)) + 
  
  geom_point() +
  geom_line() +
  
  # Error bars
  geom_errorbar(limits_conc_soil, width=.1, position=pd) +
  # scale_y_continuous(limits=c(0,10),oob = rescale_none) +
  
  # Themes and axes
  theme_minimal() +
  theme(legend.position = "right",
        # axis.text.x=element_text(angle = 45, hjust = 1),
        # axis.text.x=element_blank(), 
        axis.title.x=element_blank()
        ) +

  ylab(expression(atop("Conc. S-met ", paste({({mu}*g / g.soil.dry)} )))) +
  # facet_wrap(~Transect, nrow = 3) +
  # xlab("Date") +
  # theme() +
  # scale_x_datetime(breaks = date_breaks("1 weeks"), labels = date_format("%b %d")) +
  
  # Smooth linear models
  # stat_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  # stat_smooth(method = "lm") +
  
  # Text
  # W0 Application
  # annotate("text", x = as.POSIXct('2016-03-25 08:04:00'), y = 4, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), 
                   y = 0.5, 
                   xend = as.POSIXct('2016-03-26 01:04:00'), yend = -0), color = "black",
               arrow = arrow(length = unit(0.2, "cm"))) +

  # W1 Application
  #geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), y = 0.5, 
  #                 xend = as.POSIXct('2016-04-05 08:04:00'), yend = 0), color = "black", 
  #             arrow = arrow(length = unit(0.2, "cm"))) +
  # W2 Application
  annotate("text", x = as.POSIXct('2016-04-15 08:04:00'), y = 1, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-04-14 08:04:00'), y = 0.5, 
                   xend = as.POSIXct('2016-04-13 08:04:00'), yend = 0), color = "black", 
               arrow = arrow(length = unit(0.2, "cm"))) +
  # W9 Application
  annotate("text", x = as.POSIXct('2016-05-26 08:04:00'), y = 4.5, label = lb1a2, parse = T, size = 3.0) +
  geom_segment(aes(x = as.POSIXct('2016-05-26 08:04:00'), y = 4, 
                   xend = as.POSIXct('2016-05-26 08:04:00'), yend = 0), color = "black", 
               arrow = arrow(length = unit(0.2, "cm"))) +
  guides(col = guide_legend(nrow = 3, title.position = "top")) +
  scale_x_datetime(limits = lims, breaks = date_breaks("2 weeks"), labels = date_format("%b %d"))

if (tags){
  co + geom_text_repel(aes(label=as.factor(Wnum)),
                 size = 3,
                  arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                  force = 0.5, 
                  point.padding = unit(0.5, 'lines'), 
                 max.iter = 2e3,
                nudge_x = .05,
                show.legend = F)
}


co

# Linear model
# ggsave(co, filename = "CompositeConcLM.png", width = 7, height = 5, units = "in", scale = 1)

## ggsave(co, filename = "CompositeConcLM.tiff", height = 10, width = 8.7, units = 'cm')

# No linear model
# ggsave(co, filename = "CompositeConc.png", width = 7, height = 5, units = "in", scale = 1)

```

## Draw D\*/B\* factors

```{r}
moltenSoil = read.csv2("Data/moltenDB_R.csv", header = TRUE, sep = ",", dec = ".")
moltenSoil$Date = as.POSIXct(moltenSoil$Date, "%Y-%m-%d %H:%M", tz = "EST")

moltenSoil$DB_max = ifelse(moltenSoil$DB_max < 0, NA, moltenSoil$DB_max) 
moltenSoil$DB_max = ifelse(moltenSoil$DB_max > 1, NA, moltenSoil$DB_max) 
mSoils = melt(moltenSoil, id=c("Date", "Transect"))

mSoils$Transect = as.character(mSoils$Transect)
mSoils$Transect = ifelse(mSoils$Transect == "N", "North", 
                             ifelse(mSoils$Transect == "T", "Valley",
                                    ifelse(mSoils$Transect == "S", "South", NA)
                             )
)
mSoils$Transect <- factor(mSoils$Transect, levels = c("North", "Valley", "South"))
levels( mSoils$Transect)

DBLeg = ggplot(data = mSoils , aes(x=Date, y=value, shape=variable))+
        geom_point(data=subset(mSoils, variable == 'DB_lab')) +
        geom_point(data=subset(mSoils, variable == 'DB_max')) +
        theme_minimal() +
        guides(shape=guide_legend(title=("D*/B*"), 
                                  order = 2,
                                  nrow=2, 
                                  title.position = "top" #, 
                                  # keyheight = 1.5, 
                                  # title.vjust = NULL, label.vjust = NULL
                                  )) +
        ylab(expression(atop("D*/B*"))) +
        # facet_wrap(~Transect, nrow = 3) +
        xlab("Date") +
        scale_x_datetime(limits = lims, breaks = date_breaks("2 weeks"), labels = date_format("%b %d"))

db_leg = get_legend(DBLeg)

DB = ggplot(data = mSoils , aes(x=Date, y=value, colour=Transect, shape=variable))+
        geom_line(data=subset(mSoils, variable == 'DB_lab')) +
        geom_point(data=subset(mSoils, variable == 'DB_lab')) +
        geom_point(data=subset(mSoils, variable == 'DB_max')) +
        theme_minimal() +
        guides(shape=guide_legend(title=("D*/B* Scenario"), 
                                  order = 2,
                                  nrow=2, 
                                  title.position = "top" #, 
                                  # keyheight = 1.5, 
                                  # title.vjust = NULL, label.vjust = NULL
                                  )) +
        ylab(expression(atop("D*/B*"))) +
        # facet_wrap(~Transect, nrow = 3) +
        xlab("Date") +
        scale_x_datetime(limits = lims, breaks = date_breaks("2 weeks"), labels = date_format("%b %d"))

DB
```



## Merge hyetograph, outlet and soil concentrations

```{r}

hy_noLeg <- hy + theme(legend.position = 'none')
out_noLeg = out +  theme(legend.position = 'none')
co_noLeg = co +  theme(legend.position = 'none')

out_leg <- get_legend(out)
co_leg <- get_legend(co)


grid1 = plot_grid(
      hy_noLeg, 
      out_noLeg,
      co_noLeg,
      ncol = 1,
      align = 'hv')

ml = ggdraw() +
  draw_plot(grid1, x=0.02, y = 0.0, width = 0.95, height = .95) +
  draw_plot(out_leg, x=(0.53 ), y = 0.44, width = .8, height = 0.1) +
  draw_plot(co_leg, x=(0.53) , y = 0.11, width = .8, height = 0.1) +
  
  draw_label("A", x= 0.94, y = .93, size = 12, fontface = "bold") +
  draw_label("B", x= 0.94, y = .62, size = 12, fontface = "bold") +
  draw_label("C", x= 0.94, y = .31, size = 12, fontface = "bold")
ml





SAVE = F
PC = F
if (SAVE){
  if (PC){
      ggsave(ml, 
           filename = "D:/Documents/these_pablo/WriteUp/WaterResearch/images/multilevel.pdf", 
           device = "pdf", dpi = 600,  scale = 1, # )# ,
           width = 7, height = 11)
    } else {
      ggsave(ml, 
         filename = "/Users/DayTightChunks/Documents/PhD/Writeups/Journals/WaterResearch/images/multilevel.pdf", 
         device="pdf", dpi = 600,  scale = 1, # )# ,
         width = 7, height = 11)
    }
  }

```



# Grid arrange

```{r}


#ggsave(balAllplot, filename = "images/MB_CSIA_Bars.png", width = 8, height = 5, units = "in", scale = 1)
## convert plots to gtable objects
library(gtable)
library(grid) # low-level grid functions are required
g1 <- ggplotGrob(hy_noLeg)
# g1 <- gtable_add_cols(g1, unit(0,"mm")) # add a column for missing legend
g2 <- ggplotGrob(out_noLeg)
g3 <- ggplotGrob(co_noLeg)
# g <- rbind(g2, g3, size="first") # stack the two plots
g <- rbind(g1, g2, g3, size="first") # stack the two plots
g$widths <- unit.pmax(g1$widths, g2$widths, g3$widths) # use the largest widths
# center the legend vertically
# g$layout[grepl("guide", g$layout$name),c("t","b")] <- c(1,nrow(g))
grid.newpage()
grid.draw(g)


# Plot grid option, embedding different plot_grid 's

library(cowplot)

theme_set(theme_minimal())

embeds <- 
  plot_grid(
    plot_grid(
      hy_noLeg, 
      out_noLeg,
      co_noLeg,
      ncol = 1,
      align = 'hv'),
    plot_grid(
      ggplot(),
      out_leg,
      co_leg,
      ncol =1
    ), rel_widths = c(9,1)
  )



  
  
```


# Sources (Plot with map)

```{r}
# Water and soils transposed data
WaterSoils <- read.csv2("Data/WaterSoils_R.csv")
WaterSoils$Date.ti = as.POSIXct(strptime(WaterSoils$Date.ti,
                                 "%Y-%m-%d %H:%M", tz="EST") )
sum(is.na(WaterSoils$Date.ti))
# View(WaterSoils)
# Concnetrations to plot as box plot
keepDDtest <- c(
  "Date.ti",
  "diss.d13C.x", # "DD13C.diss",
  "comp.d13C.North", "comp.d13C.Talweg", "comp.d13C.South" #,
  #"DD13C.North", "DD13C.Talweg", "DD13C.South"
)

ws <- WaterSoils[, colnames(WaterSoils) %in% keepDDtest]
ws <- melt(ws, id="Date.ti")

ws$Source <- ifelse(ws$variable == "diss.d13C.x" , "Outlet",
                    ifelse(ws$variable == "comp.d13C.South", "South",
                           ifelse(ws$variable == "comp.d13C.Talweg", "Valley", 
                                  ifelse(ws$variable == "comp.d13C.North", "North", NA 
                    ))))

ws$Season = ifelse(ws$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Late", "Early")

ws$Group4 <- ifelse(ws$variable == "diss.d13C.x" & # Outlet
                                   ws$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Outlet(Late)",
                                 ifelse(ws$variable == "diss.d13C.x" & 
                                   ws$Date.ti <= as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Outlet(Early)",
                                 # South
                                 ifelse(ws$variable == "comp.d13C.South" & 
                                   ws$Date.ti < as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "South(Early)",
                                   ifelse(ws$variable == "comp.d13C.South" & 
                                     ws$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "South(Late)",
                                # Valley
                                     ifelse(ws$variable == "comp.d13C.Talweg" & 
                                   ws$Date.ti < as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Valley(Early)",
                                   ifelse(ws$variable == "comp.d13C.Talweg" & 
                                     ws$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Valley(Late)",
                                 # North
                                     ifelse(ws$variable == "comp.d13C.North" & 
                                   ws$Date.ti < as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "North(Early)",
                                   ifelse(ws$variable == "comp.d13C.North" & 
                                     ws$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "North(Late)", NA
                                  ))))))))

ws$Source = factor(ws$Source, levels = c( "Outlet",  "Valley", "North", "South"))
levels(ws$Source)
ws$Source = as.factor(ws$Source)

sources <- ggplot(ws, aes(Source, value)) + 
  geom_boxplot(aes(colour=Season)) + 
  theme_minimal() +
  
  # guides(col = guide_legend(nrow = 2)) +
  #theme(axis.text.x=element_text(angle = 45, hjust = 1) ,
        #axis.text.x=element_blank(),
        #axis.title.x=element_blank(),
        # legend.position="bottom"
  #      )+
  # ylab(expression(paste('S-met ', {delta}^"13","C", ' (\u2030)'))) 
  ylab(expression(atop(paste({delta}^"13","C", '  S-met '), paste({' (\u2030)'} )))) 

sources_big = sources + theme(text = element_text(size=17))

sources
  
late = subset(ws, Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'))
early = subset(ws, Date.ti <= as.POSIXct('2016-05-25 00:06:00', tz = 'EST'))
  
# Non-parameteric
# Reject Ho that pop. means are the same if low p-value
earlyGr4 <- kruskal.test(value ~ as.factor(Group4), data = early) 
earlyGr4 # result is high p value, thus cannot reject that pop. are the same (therefore, no difference)

lateGr4 <- kruskal.test(value ~ as.factor(Group4), data = late)
lateGr4 # Low p-value, therefore reject Ho that groups are the same.

SAVE = F
PC = F
if (SAVE){
  if (PC){
      ggsave(sources, 
           filename = "D:/Documents/these_pablo/WriteUp/WaterResearch/images/boxplot.pdf", 
           device = "pdf", dpi = 600,  scale = 1, # )# ,
           width = 8.7, height = 4)
    } else {
      ggsave(sources, 
         filename = "/Users/DayTightChunks/Documents/PhD/Writeups/Journals/WaterResearch/images/boxplot.pdf", 
         device=cairo_pdf, dpi = 600,  scale = 1, # )# ,
         width = 8.7, height = 4)
    }
}

```
## Grid with sources

```{r}

out_dates_noLeg = out_dates  + theme(legend.position = 'none')

sources_noLeg = sources + theme(legend.position = 'none')
so_leg = get_legend(sources) 

grid2 = plot_grid(
      hy_noLeg, 
      out_dates_noLeg,
      sources_noLeg,
      ncol = 1,
      align = 'v')

ml2 = ggdraw() +
  draw_plot(grid2, x=0.02, y = 0.0, width = 0.95, height = .95) +
  draw_plot(out_leg, x=(0.53 ), y = 0.44, width = .8, height = 0.1) +
  draw_plot(so_leg, x=(0.53) , y = 0.11, width = .8, height = 0.1) +
  
  draw_label("A", x= 0.94, y = .93, size = 12, fontface = "bold") +
  draw_label("B", x= 0.94, y = .62, size = 12, fontface = "bold") +
  draw_label("C", x= 0.94, y = .31, size = 12, fontface = "bold")
ml2
```

## All four graphs together

```{r}

hy_noLeg <- hy + theme(legend.position = 'none')
out_noLeg = out +  theme(legend.position = 'none')
co_noLeg = co +  theme(legend.position = 'none')

out_leg <- get_legend(out)
co_leg <- get_legend(co)

db_noLeg = DB + theme(legend.position = 'none')

grid3 = plot_grid(
      hy_noLeg, 
      out_noLeg,
      co_noLeg,
      db_noLeg,
      # sources_noLeg,
      ncol = 1,
      align = 'v')

ml3 = ggdraw() +
  draw_plot(grid3, x=0.02, y = 0.0, width = 0.95, height = .95) +
  draw_plot(out_leg, x=(0.54 ), y = 0.56, width = .8, height = 0.1) +
  draw_plot(co_leg, x=(0.54 ), y = 0.32, width = .8, height = 0.1) +
  draw_plot(db_leg, x=(0.54) , y = 0.08, width = .8, height = 0.1) +
  
  draw_label("A", x= 0.94, y = .93, size = 12, fontface = "bold") +
  draw_label("B", x= 0.94, y = .69, size = 12, fontface = "bold") +
  draw_label("C", x= 0.94, y = .47, size = 12, fontface = "bold") +
  draw_label("D", x= 0.94, y = .23, size = 12, fontface = "bold")
ml3
```
## Save

```{r}
SAVE = F
PC = F
if (SAVE){
  if (PC){
      ggsave(ml3, 
           filename = "D:/Documents/these_pablo/WriteUp/WaterResearch/images/multilevel4.pdf", 
           device = "pdf", dpi = 600,  scale = 1, # )# ,
           width = 7, height = 10)
    } else {
      ggsave(ml3, 
         filename = "/Users/DayTightChunks/Documents/PhD/Writeups/Journals/WaterResearch/images/multilevel4.pdf", 
         device=cairo_pdf, dpi = 600,  scale = 1, # )# ,
         width = 7, height = 10)
    }
  }
```


## Double isotope graph

```{r}

oDual = WaterSoils[ , c("Date.ti", "WeekSubWeek", "diss.d13C.x", "SD.d13C.x")]
oDual$WeekSubWeek = as.character(oDual$WeekSubWeek)
  
oDual$WeekSubWeek = ifelse(oDual$Date.ti == as.POSIXct("2016-03-25 00:04:00", tz= "EST"), "Product", oDual$WeekSubWeek)
oDual$diss.d13C.x = ifelse(oDual$Date.ti == as.POSIXct("2016-03-25 00:04:00", tz= "EST"), -32.2, oDual$diss.d13C.x)
oDual$SD.d13C.x = ifelse(oDual$Date.ti == as.POSIXct("2016-03-25 00:04:00", tz= "EST"), 0.4, oDual$SD.d13C.x)


n = read.csv2("Data/NitrogenIRMS.csv", dec = ".")
n$WeekSubWeek = as.character(n$WeekSubWeek)
n[nrow(n)+1,] = c("Product", as.numeric(1.9), as.numeric(0.5))

#row = c("Product", as.numeric(1.9), as.numeric(0.5)) 
#n = rbind(n, row)
#str(n)
oDual = merge(oDual, n, by = "WeekSubWeek", all = F)
oDual$diss.d15N = as.numeric(oDual$diss.d15N)
oDual$SD.d15N = as.numeric(oDual$SD.d15N)

oDual$Month = ifelse(oDual$Date.ti >= as.POSIXct("2016-03-24 00:30:00", tz = "EST") &
           oDual$Date.ti < as.POSIXct("2016-04-01 00:00:00", tz = "EST"), "April",
         ifelse(oDual$Date.ti >= as.POSIXct("2016-04-01 00:00:00", tz = "EST") &
                  oDual$Date.ti < as.POSIXct("2016-05-01 00:00:00", tz = "EST"), "April",
                ifelse(oDual$Date.ti >= as.POSIXct("2016-05-01 00:00:00", tz = "EST") &
                         oDual$Date.ti < as.POSIXct("2016-06-01 00:00:00", tz = "EST"), "May",
                       ifelse(oDual$Date.ti >= as.POSIXct("2016-06-01 00:00:00", tz = "EST") & 
                                oDual$Date.ti < as.POSIXct("2016-07-01 00:00:00", tz = "EST"), "June", "July" )
                            )
                     )
         )

oDual$Season = ifelse(oDual$Date.ti > as.POSIXct('2016-05-25 00:06:00', tz = 'EST'), "Late", "Early")

y = oDual$diss.d15N
y.sd = oDual$SD.d15N

x = oDual$diss.d13C.x
x.sd = oDual$SD.d13C.x

d = ggplot(data = oDual, aes(x=x, y=y, group = Month, colour = Month, shape = Month))+
  theme_minimal() +
  geom_errorbar(aes(ymin = y - y.sd, ymax = y + y.sd)) +
  geom_errorbarh(aes(xmin = x - x.sd, xmax = x + x.sd)) +
  geom_point() +
  ylab(expression(paste({delta}^"15","N", ' (\u2030)'))) +
  xlab(expression(paste({delta}^"13","C", ' (\u2030)'))) +
  theme(text = element_text(size=17))
  
# View(subset(oDual, WeekSubWeek == "Product" ))

tags = T
oDual$Label = ifelse(oDual$WeekSubWeek != "Product", NA, "Product") #  as.factor(oDual$WeekSubWeek)
if (tags){
d = d +  geom_label_repel(data = subset(oDual),  aes(label=Label),
                 arrow = arrow(length = unit(0.005, 'npc'), type = "closed"),
                 force = 0.1, 
                 point.padding = unit(1, 'lines'), 
                 max.iter = 2e3,
                 nudge_x = .1, nudge_y = -0.2, show.legend = F)
}

d = d + scale_x_continuous(breaks = seq(-33, -29, by=0.5) )

d

SAVE = F
PC = T
if (SAVE){
  if (PC){
      ggsave(d, 
           filename = "D:/Documents/these_pablo/WriteUp/WaterResearch/images/dual.pdf", 
           device = "pdf", dpi = 600,  scale = 1, # )# ,
           width = 7, height = 6)
    } else {
      ggsave(d, 
         filename = "/Users/DayTightChunks/Documents/PhD/Writeups/Journals/WaterResearch/images/dual.pdf", 
         device=cairo_pdf, dpi = 600,  scale = 1, # )# ,
         width = 7, height = 6)
    }
  }

```

